---
title: "Birmingham_diag_explore"
output: html_document
date: "2025-02-24"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE,message = FALSE)
library(MASS)
library(tidyverse)
library(latex2exp)
library(viridis)
#library(evd)
#library(plgp)
#library(gridExtra)
library(here)
library(tmap)
#library(units)
#library(RColorBrewer)
library(gnorm)
library(sf)
file.sources = list.files(pattern="*helpers.R")
sapply(file.sources,source,.GlobalEnv)
source("rotate_unrotate_coordinates.R")
gr_npole_lat <- 37.5
gr_npole_lon <- 177.5
load("spatialobjects.RData") # uk polygon and uk_temp_sf points sf objects
load("est_all.RData") # for est_all_diag R object
theme_set(theme_bw())
theme_replace(
  panel.spacing = unit(2, "lines"),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  strip.background = element_blank(),
  panel.border = element_rect(colour = "black", fill = NA) )

ukcp18 <- readRDS("data/uk_1999_2018_summer.RDS") %>% relocate(dist_london,.after=dist_glasgow)
# remove last year of the data due to error (same data as first year)
ukcp18 <- ukcp18[,1:1716]

conv <- CnvRttPol(latlon = data.frame(long=ukcp18$Longitude,lat=ukcp18$Latitude),spol_coor = c(gr_npole_lon, gr_npole_lat))
uk_sf_rot <- data.frame(lon=conv$lon,lat=conv$lat,ind=conv$ind) %>%
  st_as_sf(coords = c("lon", "lat"), crs = 4326)
uk_sf_rot <- st_transform(uk_sf_rot,27700) %>% mutate(lon=conv$lon,lat=conv$lat)
sims <- ukcp18 %>% dplyr::select(!contains("i")) %>% t() %>% as.data.frame()
# ordered alphabetically so Y1 Birmingham, Y2 Glasgow and Y3 is London
colnames(sims) <- paste0("Y",1:ncol(sims))
# transform to Laplace margins
sims <- as.data.frame((sims %>% apply(c(2),FUN=row_number))/(nrow(sims)+1)) %>% apply(c(1,2),FUN=unif_laplace_pit) %>% as.data.frame()
v <- 0.9 # set threshold

```

## Plot sites on the diagonal from Birmingham to London

```{r}
sites_index_diagonal <- c(161,144,128,110,92,77,78) # first is Birmingham and last is London
site_name_diagonal <- c("Birmingham", paste0("diagonal",1:5),"London")
# plot these points on a map
uk_sf_rot <- uk_sf_rot %>% mutate(sites_index_diagonal=factor(case_match(ind,c(161) ~ "Birmingham",c(78)~"London",sites_index_diagonal[2:6]~"diagonal_sites")))
tm_shape(uk_sf_rot) + tm_dots("sites_index_diagonal",size=0.1, palette = c("Birmingham"="#C11432","London" = "#009ADA", "diagonal_sites" = "#FDD10A"))
```
Estimate all parameters for the sites in colour (including Birmingham and London for testing).

```{r, echo=FALSE}
dayshift <- c(-3:3)
# est_all_diag <- data.frame("lik" = numeric(), "lika"= numeric(),"likb"= numeric(),"lik2"= numeric(),
#                           "a" = numeric(), "b" = numeric(),
#                           "mu" = numeric(),"mu_agg" = numeric(),
#                           "sig" = numeric(),"sig_agg" = numeric(),"sigl" = numeric(),"sigu" = numeric(),
#                           "delta" = numeric(),"deltal" = numeric(), "deltau" = numeric(),
#                           "given" = numeric(), "res" = numeric(), "cond_site" = character(), "tau" = numeric())
# for (i in 1:length(sites_index_diagonal)) {
#   cond_site <- sites_index_diagonal[i]
#   for (j in 1:length(dayshift)) {
#     sims_tau <- shift_time(sims=sims,cond_site=cond_site,tau=dayshift[j])
#     est_all_diag <- rbind(est_all_diag,par_est(df=sims_tau,v=v,given=cond_site,margin = "AGG", method="sequential2",keef_constraints = c(1,2)) %>% add_row(.before=cond_site) %>%  mutate(cond_site=sites_name_diagonal[i],tau=as.character(dayshift[j])))
#   }
# }
```

Plot summary of parameters for the sites on the diagonal.

```{r}
est_all_diag1 <- est_all_diag %>% mutate(cond_site = factor(cond_site,levels=sites_index_diagonal), tau = factor(tau,levels=as.character(dayshift)))
est_all_diag1 %>% ggplot() + geom_boxplot(aes(x=cond_site, y=a))
est_all_diag1 %>% ggplot() + geom_boxplot(aes(x=cond_site, y=b))
```

There is a decrease in both $\hat{\boldsymbol{\alpha}}$ and $\hat{\boldsymbol{\beta}}$.

```{r}
est_all_diag1 %>% group_by(cond_site) %>% summarise(mean_a=mean(a,na.rm = TRUE),mean_b = mean(b, na.rm = TRUE),median_a=median(a,na.rm = TRUE),median_b = median(b, na.rm = TRUE))
```
This decrease is non-linear and can be observed in both mean and median values of $\hat{\boldsymbol{\alpha}}$ and $\hat{\boldsymbol{\beta}}$.

### Map parameter estimates 

```{r}
condmodel_params <- c("a","b","mu","sig","muagg","sigl","sigu","sigdiff","deltal","deltau","deltadiff")
# for (i in 1:length(sites_index_diagonal)) {
# tau_site_map <- map_param(tmp_est=est_all_diag1 %>% filter(cond_site== sites_index_diagonal[i]), method = "AGG", facet_var = "tau", title_map = site_name_diagonal[i],grid_uk = uk_temp_sf)
#   for (j in 1:length(condmodel_params)) {
#     tmap_save(tau_site_map[[j]],filename=paste0("../Documents/est_all_diag/tau_",site_name_diagonal[i],"_map_",j,"_",condmodel_params[j],".png"),width=15,height=4.5)
#   }
# }

# map also a grid of alpha and beta estimates
est_all_diag1 <- est_all_diag1 %>% mutate(cond_site = factor(as.character(cond_site),levels=as.character(sites_index_diagonal)))
tau_map <- map_param(tmp_est=est_all_diag1, method = "AGG", facet_var = c("cond_site","tau"), title_map = "Diagonal sites",grid_uk = uk_temp_sf)
tau_map[[1]]
tau_map[[2]]
  # for (j in 1:length(condmodel_params)) {
  #   tmap_save(tau_map[[j]],filename=paste0("../Documents/est_all_diag/allsites_tau_map_",j,"_",condmodel_params[j],".png"),width=6,height=10)
  # }
```

Plot also return levels for conditional quantiles as before.

```{r, echo=FALSE}
tmp <- est_all_diag1
# x0999 <- unif_laplace_pit(0.999)
# q025 <- q075 <- numeric()
# for (i in 1:nrow(tmp)) {
#     # filter value of a and b
#   a <- as.numeric(tmp[i,5])
#   b <- as.numeric(tmp[i,6])
#   if (is.na(a)) {q025[i] <- q075[i] <- NA} else {
#   # find residual variable and cond.variable
#   cond_site. <- as.numeric(as.character(tmp$cond_site[i]))
#   res_site <- tmp$res[i]
#   # calculate Y1 and Yj
#   sims_tau <- shift_time(sims=sims,cond_site=cond_site.,tau=as.numeric(as.character(tmp$tau[i]))) %>% dplyr::select(all_of(c(cond_site.,res_site)))
#   names(sims_tau) <- c("Y1","Y2")
#   Y1_Y2 <- sims_tau %>% filter(sims_tau[,1]>quantile(sims_tau[,1],v))
#   Y1 <- as.numeric(Y1_Y2[,1])
#   Y2 <- as.numeric(Y1_Y2[,2])
#   # calculate observed residuals
#   z <- (Y2-a*Y1)/(Y1^b)
#   # calculate q025 and q075
#   q025[i] <- a*x0999+x0999^b*quantile(z,0.25)
#   q075[i] <- a*x0999+x0999^b*quantile(z,0.75)
#   }
# }
# 
# # join and map
# tmp <- tmp %>% mutate(q025,q075) %>%  pivot_longer(,cols=c(q025,q075),names_to = "q") %>% arrange(q) %>% mutate(q=factor(q)) %>% mutate(tau=factor(as.character(tau),levels=unique(as.character(tmp$tau))))
# 
# rl025 <- sapply(1:length(q025),function(i){1/(1-pgnorm(q=q025[i]))})
# rl075 <- sapply(1:length(q075),function(i){1/(1-pgnorm(q=q075[i]))})
# tmp <- tmp %>% mutate(rl=c(rl025,rl075))
# tmp025 <- tmp %>% filter(q=="q025")
# tmp075 <- tmp %>% filter(q=="q075")
# tmpdiff <- tmp %>% filter(q=="q025") %>% mutate(rldiff=(rl075-rl025),rlreldiff=(rl075-rl025)/rl075) %>%  dplyr::select(cond_site,tau,rldiff,rlreldiff,given,res)

tmp025_q <- map_param(tmp_est=tmp025,method = "q", facet_var = c("cond_site","tau"),title_map = TeX("$p=0.25$"))
tmp025_rl <- map_param(tmp_est=tmp025,method = "rl", facet_var = c("cond_site","tau"),title_map = TeX("$p=0.25$"))
# tmap_save(tmp025_q,filename=paste0("../Documents/est_all_diag/q025_tau.png"),width=6,height=10)
# tmap_save(tmp025_rl,filename=paste0("../Documents/est_all_diag/rl025_tau.png"),width=6,height=10)

tmp075_q <- map_param(tmp_est=tmp075,method = "q", facet_var = c("cond_site","tau"),title_map = TeX("$p=0.75$"))
tmp075_rl <- map_param(tmp_est=tmp075,method = "rl", facet_var = c("cond_site","tau"),title_map = TeX("$p=0.75$"))
# tmap_save(tmp075_q,filename=paste0("../Documents/est_all_diag/q075_tau.png"),width=6,height=10)
# tmap_save(tmp075_q,filename=paste0("../Documents/est_all_diag/rl075_tau.png"),width=6,height=10)

# plot return level relative difference
tmap_reldiff <- map_param(tmp_est=tmpdiff,method = "rlreldiff", facet_var = c("cond_site","tau"),title_map = TeX(""))
# tmap_save(tmap_reldiff,filename=paste0("../Documents/est_all_diag/rlreldiff_tau.png"),width=6,height=10)

# plot also relative difference
# plot return level difference
tmap_diff <- map_param(tmp_est=tmpdiff,method = "rldiff", facet_var = c("cond_site","tau"),title_map = TeX(""))
# tmap_save(tmap_diff,filename=paste0("../Documents/est_all_diag/rldiffrel_tau.png"),width=6,height=10)
```

```{r}
tmp025_q
tmp075_q
```

And also plot return levels.

```{r}
tmp025_rl
tmp075_rl
```
And return level difference.
```{r}
tmap_diff
```

And relative return level difference.

```{r}
tmap_reldiff
```

