---
title: "Compare_residual_margins"
output: html_document
date: "2024-09-24"
---

```{r setup, include=FALSE}
library(tidyverse)
library(latex2exp)
library(gridExtra)
library(gnorm)
knitr::opts_chunk$set(echo = TRUE)
# set theme defaults to be black rectangle
theme_set(theme_bw())
theme_replace(
  panel.spacing = unit(2, "lines"),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  strip.background = element_blank(),
  panel.border = element_rect(colour = "black", fill = NA) )
file.sources = list.files(pattern="*helpers.R")
sapply(file.sources,source,.GlobalEnv)
```

## Simulate data

Simulate from trivariate Markov chain with logistic copula links and Laplace margins.

```{r }
set.seed(11)
N <- 500
v <- 0.99 # threshold for conditioning variable
sims <- generate_Y(N=N) %>% link_log(dep=1/2) %>%
  link_log(dep=1/2) %>% link_log(dep=1/2) %>% link_log(dep=1/2) %>%
  apply(c(1,2),FUN=frechet_laplace_pit) %>% as.data.frame()
```

## Motivate the problem

Simulate a $n=50000$ sample from a logistic distribution.

```{r}
set.seed(11)
N <- 50000
v <- 0.99 # threshold for conditioning variable
sims <- generate_Y(N=N) %>% link_log(dep=1/2) %>%
  link_log(dep=1/2) %>% link_log(dep=1/2) %>% link_log(dep=1/2) %>%
  apply(c(1,2),FUN=frechet_laplace_pit) %>% as.data.frame()
obsr <- observed_residuals(df=sims,given=1,v=v)
df <- data.frame(y=g_laplace(z=seq(-7.5,5,0.01),a=1/2),x=seq(-7.5,5,0.01))
ggplot(obsr) + geom_density(mapping = aes(x=Z2)) + geom_line(data=df,mapping = aes(x=x,y=y),col="#C11432") + xlab(TeX("$Z_2$")) + ylab("Residual density function")
```

## Calculate MLE for different methods

```{r}
L1 <- par_est(df=sims,v=v,given=c(1:5),margin = "GenGaus", method = "two_step")
L2 <- par_est(df=sims,v=v,given=c(1:5),margin = "AGGsig", method = "two_step")
L3 <- par_est(df=sims,v=v,given=c(1:5),margin = "AGG", method = "two_step")
L4 <- par_est(df=sims,v=v,given=c(1:5),margin = "AGGsigdelta", method = "two_step")

sum(2*((L2$lik2)-(L1$lik2))>pchisq(0.95,1))/nrow(L1)
sum(2*((L3$lik2)-(L1$lik2))>pchisq(0.99,1))/nrow(L1)
sum(2*((L4$lik2)-(L1$lik2))>pchisq(0.99,2))/nrow(L1)

2*((L4$lik2)-(L1$lik2))>pchisq(0.95,2)
2*((L3$lik2)-(L2$lik2))
2*((L3$lik2)-(L1$lik2))

```

Specifically compare considering different scale or shape parameter or both.
```{r}
N <- 50000
v <- 0.99
M <- 100
load("paramest.rda")
# L1 <- list()
# L2 <- list()
# L3 <- list()
# L4 <- list()
# for (i in 1:M) {
#   # simulate a sample
#   set.seed(12*i)
#   sims <- generate_Y(N=N) %>% link_log(dep=1/2) %>%
#     link_log(dep=1/2) %>% link_log(dep=1/2) %>% link_log(dep=1/2) %>%
#     apply(c(1,2),FUN=frechet_laplace_pit) %>% as.data.frame()
# # calculate log-likelihoods cond on 1 and save in a list
# L1[[i]] <- par_est(df=sims,v=v,given=c(1),margin = "GenGaus", method = "two_step")
# L2[[i]] <- par_est(df=sims,v=v,given=c(1),margin = "AGGsig", method = "two_step")
# L3[[i]] <- par_est(df=sims,v=v,given=c(1),margin = "AGG", method = "two_step")
# L4[[i]] <- par_est(df=sims,v=v,given=c(1),margin = "AGGsigdelta", method = "two_step")
# }
load("paramest.rda")

# plot boxplots for likelihood
ablik <- c()
GenGaus <- c()
AGGsig <- c()
AGGdelta <- c()
AGGsigdelta <- c()
res <- rep(c("Z2","Z3","Z4","Z5"),length(L1))
for (i in 1:length(L1)) {
  ablik <- append(ablik,-as.numeric(unlist(L1[[i]]$lik)))
  GenGaus <- append(GenGaus,as.numeric(unlist(L1[[i]]$lik2)))
  AGGsig <- append(AGGsig,as.numeric(unlist(L2[[i]]$lik2)))  
  AGGdelta <- append(AGGdelta,as.numeric(unlist(L3[[i]]$lik2)))
  AGGsigdelta <- append(AGGsigdelta,as.numeric(unlist(L4[[i]]$lik2)))
}
lims <- c(min(AGGdelta,AGGsig,AGGsigdelta,GenGaus),max(AGGdelta,AGGsig,AGGsigdelta,GenGaus))
p1 <- ggplot(data.frame(x=factor(res,levels=c("Z2","Z3","Z4","Z5")),y=GenGaus),aes(x=x,y=y))+
  geom_boxplot() + xlab("Observed residuals") + ylab(TeX("$l(\\mu,\\sigma,\\delta)$"))+ ylim(lims)
p2 <- ggplot(data.frame(x=factor(res,levels=c("Z2","Z3","Z4","Z5")),y=AGGsig),aes(x=x,y=y))+
  geom_boxplot() + xlab("Observed residuals") + ylab(TeX("$l(\\mu,\\sigma_l,\\sigma_u,\\delta)$")) + ylim(lims)
p3 <- ggplot(data.frame(x=factor(res,levels=c("Z2","Z3","Z4","Z5")),y=AGGdelta),aes(x=x,y=y))+
  geom_boxplot() + xlab("Observed residuals") + ylab(TeX("$l(\\mu,\\sigma,\\delta_l,\\delta_u)$"))+ ylim(lims)
p4 <- ggplot(data.frame(x=factor(res,levels=c("Z2","Z3","Z4","Z5")),y=AGGsigdelta),aes(x=x,y=y))+
  geom_boxplot() + xlab("Observed residuals") + ylab(TeX("$l(\\mu,\\sigma_l,\\sigma_u,\\delta_l,\\delta_u)$")) + ylim(lims)
p5 <- ggplot(data.frame(x=factor(res,levels=c("Z2","Z3","Z4","Z5")),y=ablik),aes(x=x,y=y))+
  geom_boxplot() + xlab("Observed residuals") + ylab(TeX("$l(\\alpha,\\beta,\\mu,\\sigma)$"))+ ylim(c(min(ablik,lims[1]),max(ablik,lims[2])))

grid.arrange(p1,p2,p3,p4,ncol=2)
```

Now, plot likelihood ratio test values instead with $\chi$

```{r}
p1 <- ggplot(data.frame(x=factor(res,levels=c("Z2","Z3","Z4","Z5")),y=2*(AGGsigdelta-GenGaus)),aes(x=x,y=y))+
  geom_boxplot() + xlab("Observed residuals") +
  # ylab("Different scale vs shape")
  ylab(TeX("$2(l(\\mu,\\sigma_l,\\sigma_u,\\delta_l,\\delta_u)-l(\\mu,\\sigma,\\delta))$")) + ylim(c(0,max(2*(AGGsigdelta-GenGaus)))) +
   geom_hline(yintercept = qchisq(0.95,2),linetype="dashed",col = "#66A64F") +  geom_hline(yintercept = qchisq(0.99,2),linetype="dashed", col= "#FDD10A")
p2 <- ggplot(data.frame(x=factor(res,levels=c("Z2","Z3","Z4","Z5")),y=2*(AGGsig-GenGaus)),aes(x=x,y=y))+
  geom_boxplot() + xlab("Observed residuals") +
  # ylab("No difference in shape")
  ylab(TeX("$2(l(\\mu,\\sigma_l,\\sigma_u,\\delta)-l(\\mu,\\sigma,\\delta))$")) +
ylim(c(0,max(2*(AGGsigdelta-GenGaus)))) +
  geom_hline(yintercept = qchisq(0.95,1),linetype="dashed",col = "#66A64F") +  geom_hline(yintercept = qchisq(0.99,1),linetype="dashed", col= "#FDD10A")
p3 <- ggplot(data.frame(x=factor(res,levels=c("Z2","Z3","Z4","Z5")),y=2*(AGGdelta-GenGaus)),aes(x=x,y=y))+
  geom_boxplot() + xlab("Observed residuals") +
  # ylab("Different scale vs shape") +
  ylab(TeX("$2(l(\\mu,\\sigma,\\delta_l,\\delta_u)-l(\\mu,\\sigma,\\delta))$")) + ylim(c(0,max(2*(AGGsigdelta-GenGaus)))) +
  geom_hline(yintercept = qchisq(0.95,1),linetype="dashed",col = "#66A64F") +  geom_hline(yintercept = qchisq(0.99,1),linetype="dashed", col= "#FDD10A")
p4 <- ggplot(data.frame(x=factor(res,levels=c("Z2","Z3","Z4","Z5")),y=2*(AGGsig-AGGdelta)),aes(x=x,y=y))+
  geom_boxplot() + xlab("Observed residuals") +
  # ylab("Different scale vs shape") +
  ylab(TeX("$2(l(\\mu,\\sigma_l,\\sigma_u,\\delta)-l(\\mu,\\sigma,\\delta_l,\\delta_u))$")) 

grid.arrange(p1,p2,p3,p4,ncol=2)
```

Compare $\hat{\alpha}$ and $\hat{\beta}$.

```{r}
a <- b <- c()
for (i in 1:length(L1)) {
 a <- append(a,as.numeric(unlist(L1[[i]]$a)))
 b <- append(b,as.numeric(unlist(L2[[i]]$b)))  
}

p1 <- ggplot(data.frame(x=factor(res,levels=c("Z2","Z3","Z4","Z5")),y=a),aes(x=x,y=y))+
  geom_boxplot() + xlab("Observed residuals") + ylab(TeX("$\\hat{\\alpha}$"))
p2 <- ggplot(data.frame(x=factor(res,levels=c("Z2","Z3","Z4","Z5")),y=b),aes(x=x,y=y))+
  geom_boxplot() + xlab("Observed residuals") + ylab(TeX("$\\hat{\\beta}$")) 
grid.arrange(p1,p2,ncol=2)
```

