---
title: "Extremal dependence paper"
format: html
---

The aim is to recreate the analysis and plots from *Dependence Measures for Extreme Value Analysis* paper by Stuart Coles and Jonathan Tawn (1999).

```{r setup, warning=FALSE}
#| echo: false
#| message: false
#| warning: false
library(tidyverse)
library(latex2exp)
library(gridExtra)
library(ismev)
library(VGAM)
library(latex2exp)
library(extRemes) # for precise chi bar measure at high quantiles
library(mvtnorm)
# set theme defaults to be black rectangle
theme_set(theme_bw())
theme_replace(
          panel.spacing = unit(2, "lines"),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          strip.background = element_blank(),
          panel.border = element_rect(colour = "black", fill = NA) )

```

There are $3$ datasets in the paper. First, temperature data from Oxford and Worthing examines spatial dependence of extremely high temperatures. Second, precipitation data from Oxford to examine time dependence of precipitation. Third, dependence between surge levels and tides for extreme sea levels. 

```{r}
data(wavesurge)
# plot the wavesurge data
ggplot(wavesurge) + geom_point(aes(x=wave,y=surge))+ theme_minimal() + theme(aspect.ratio = 1) + xlab("Wave height") + ylab("Surge level")

```

Using the probability integral transform (PIT), plot the quantiles.

```{r}
# create x and y quantile variables
U <- wavesurge %>% select(wave) %>% arrange(wave) %>% mutate(u=row_number()/(nrow(wavesurge)+1))
V <- wavesurge %>% select(surge) %>% arrange(surge) %>% mutate(v=row_number()/(nrow(wavesurge)+1)) 
# examine top and bottom
U %>% head()
U %>% tail()
V %>% head()
V %>% tail()
df <- wavesurge %>% left_join(U,by="wave") %>% left_join(V,by="surge")
df %>% head()
```
Now plot the values on the $U(0,1)$ scale to replicate (c) of Figure $5$.

```{r}
ggplot(df)+ geom_point(aes(x=u,y=v)) + theme(aspect.ratio = 1) + xlab("Wave height") + ylab("Surge level")
```

### Dependence measures

Now, try to replicate Figure $10$ to plot the dependence measures $\chi(u)$ and $\Bar{\chi(u)}$.We need to calculate the empirical estimate of $C(u,u)$. Each marginal distribution is estimated by its empirical distribution function. The confidence intervals are constructed using the delta method.

To illustrate the problem of dependence measure $\chi(u)$, we can plot the function for different values of correlation coefficient $\rho$ in the Gaussian dependence model.

```{r}
# for loop for diff values of rho
U <- c()
Chi <- c()
Rho <- c()
edge <- 10^(-6)
u <- seq(0+edge,1-edge,0.01)
rho_vec <- seq(-0.9,0.9,0.1)
for (i in 1:length(rho_vec)) {
rho <- rho_vec[i]
# calculate the copula using VGAM package
c_uu <- pbinormcop(q1=u,q2=u,rho=rho)
# calculate chi(u)
chi <- 2- (log(c_uu))/(log(u))
u <- u[chi>-10^6]
chi <- chi[chi>-10^6]
# subset by Frechet bounds (lower bound is sufficient)
lower_bound1 <- chi>(2-(log(2*u-1))/(log(u)))
lower_bound <- lower_bound1 %>% replace(is.na(.), TRUE)
# add items to a list
U <- append(U,u[lower_bound])
Chi <- append(Chi,chi[lower_bound])
Rho <- append(Rho,rep(rho_vec[i],sum(lower_bound)))
}
#p <- p + geom_line(data=df,mapping = aes(x=u,y=chi))
df <- data.frame(U,Chi,Rho=as.factor(Rho))
# subset to remove \chi(u)<-1 to match Coles' paper
df <- df[Chi>=-1,]
# ggplot requires prepared data before plotting
ggplot(df) + geom_point(aes(x=U,y=Chi,color=Rho)) + xlab(TeX("$u$")) + ylab(TeX("$\\chi(u)$"))
```
Furthermore, plotting extra points close to $1$ makes the convergence more obvious.

```{r}
#| echo: false
# for loop for diff values of rho
U <- c()
Chi <- c()
Rho <- c()
edge <- 10^(-6)
u <- c(seq(0.02+edge,0.95,0.01),seq(0.95,1-edge,length.out=100))
rho_vec <- seq(-0.9,0.9,0.1)
for (i in 1:length(rho_vec)) {
rho <- rho_vec[i]
# calculate the copula using VGAM package
c_uu <- pbinormcop(q1=u,q2=u,rho=rho)
# calculate chi(u)
chi <- 2- (log(c_uu))/(log(u))
# u <- u[chi>-10^6]
# chi <- chi[chi>-10^6]
# subset by Frechet bounds (lower bound is sufficient)
lower_bound1 <- chi>(2-(log(2*u-1))/(log(u)))
lower_bound <- lower_bound1 %>% replace(is.na(.), TRUE)
# add items to a list
U <- append(U,u[lower_bound])
Chi <- append(Chi,chi[lower_bound])
Rho <- append(Rho,rep(rho_vec[i],sum(lower_bound)))
}
#p <- p + geom_line(data=df,mapping = aes(x=u,y=chi))
df <- data.frame(U,Chi,Rho=as.factor(Rho))
# subset to remove \chi(u)<-1 to match Coles' paper
df <- df[Chi>=-1,]
# ggplot requires prepared data before plotting
# ,col= guide_legend(title= TeX("$\rho$"))
p1 <- ggplot(df) + geom_line(aes(x=U,y=Chi,fill=Rho))+ xlab(TeX("$u$")) + ylab(TeX("$\\chi(u)$")) +  guides(colour = guide_legend(reverse=T)) + labs(fill=TeX("$\rho$")) + geom_segment(aes(x = 0, y = 1, xend = 1, yend = 1), linetype="dashed")
p1
```
What happens if we do not restrict lower limit of $y$-axis to be $-1$? As there is no lower bound as $u \rightarrow 0$ in the case of negative dependence, the illustration becomes harder to read.

### Dependence measure $\Bar{\chi}(u)$

Theoretical part of the paper involves plotting the dependence measures $\chi(u)$ and $\Bar{\chi(u)}$.

```{r}
# for loop for diff values of rho
U <- c()
Chi <- c()
Rho <- c()
edge <- 10^(-5)

rho_vec <- seq(0.9,-0.9,-0.1)
for (i in 1:length(rho_vec)) {
rho <- rho_vec[i]
u <- c(seq(0+edge,0.95,0.01),seq(0.95,(1-edge),length.out=50))
# calculate the copula using VGAM package
 c_uu <- c()
  #c_uu <- 1-2*u+  rvinecopulib::pbicop(u=matrix(c(u,u),ncol=2),family="gaussian",parameters=rho)
  for (i in 1:length(u)) {
  c_uu[i] <- mvtnorm::pmvnorm(lower=c(qnorm(u[i]),qnorm(u[i])),sigma=matrix(c(1,rho,rho,1),ncol=2))
      if (i>2) {if (c_uu[i]<=0) {c_uu[i] <- c_uu[i-1]} }
  }
# calculate chi(u)
 # chi <- -1+ (2*log(1-u))/(7.9999^(-1)*log(c_uu^(7.9999)))
chi <- -1+ (2*log(1-u))/(log(c_uu))
#u <- u[chi>-10^6]
#chi <- chi[chi>-10^6]
# subset by Frechet bounds (lower bound is sufficient)
#lower_bound1 <- chi>(2-(log(2*u-1))/(log(u)))
#lower_bound <- lower_bound1 %>% replace(is.na(.), TRUE)
# add items to a list
U <- append(U,u)
Chi <- append(Chi,chi)
Rho <- append(Rho,rep(rho,length(u)))
}
# rho <- -0.9
# c_uu <- c()
# chi <- c()
# c_uu <- 1-2*u+pbinormcop(q1=u,q2=u,rho=rho)
# # calculate chi(u)
# for (j in 1:length(u)) {
#   if (u[j]< 10^-10) {
#     chi[j] <- -1+ (2*log(1-u))/(log(c_uu*(10^20))-20*log(10))
#   }
#    else {
#      chi[j] <- -1+ (2*log(1-u))/log(c_uu)
#    } 
# }
#  chi <- -1+ (2*log(1-u))/log(c_uu)
#  plot(u,chi)
#p <- p + geom_line(data=df,mapping = aes(x=u,y=chi))
df <- data.frame(U,Chi,Rho=as.factor(Rho))
# subset to remove \chi(u)<-1 to match Coles' paper
df <- df[Chi>-1,] 
#df$U[(df$U==1) &(df$Rho<0.6)] <- NA
# ggplot requires prepared data before plotting
p2 <- ggplot(df %>% drop_na()) + geom_line(aes(x=U,y=Chi,fill=Rho)) + xlab(TeX("$u$")) + ylab(TeX("$\\bar{\\chi}(u)$")) + ylim(c(-1,1)) + xlim(c(0,1)) + geom_segment(aes(x = 0, y = 1, xend = 1, yend = 1), linetype="dashed") + geom_segment(aes(x = 0, y = -1, xend = 1, yend = -1),linetype="dashed")
p2
#+  guides(colour = guide_legend(reverse=T))
#grid.arrange(p1,p2,ncol=2)
```

