---
title: "Explore extremes"
format: html
---

```{r}
#| echo: false
#| message: false
#| warning: false
library(tidyverse)
library(latex2exp)
library(ismev)
library(gridExtra)
library(evd)
source("gev_fit.R")
theme_set(theme_bw())
theme_update(plot.title = element_text(hjust = 0.5),
             panel.spacing = unit(2, "lines"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        strip.background = element_blank(),
        panel.border = element_rect(colour = "black", fill = NA))
```

### Return level plots

Exercise recreating Fig. 3.1 from Coles' book. Possible typo in the book as the $x$-axis should be $- \log(-\log(1-p))$ and not $\log(-\log(1-p))$.

```{r}
# set parameters
mu <- rep(0,3)
sigma <- rep(1,3)
xi <- c(-0.2,0,0.2)
U <- runif(10000) %>% sort()
z_p_neg <- c()
z_p_pos <- c()
z_p_0 <- c()
# calculate z_p for different values of xi
for (i in seq_along(U)) {
  z_p_neg[i] <-  -(1- (-log(U[i]))^(-xi[1])  )*(sigma[1]/xi[1])+ mu[1]
  z_p_pos[i] <-  -(1- (-log(U[i]))^(-xi[3])  )*(sigma[3]/xi[3])+ mu[3]
  z_p_0[i] <- -log(-log(U[i]))*sigma[2]+mu[2]
}
z_p <- data.frame(U,z_p_neg,z_p_0,z_p_pos)
z_p %>% ggplot() + geom_line(aes(x=-log(-log(U)),y=(z_p_neg)),color="#009ada",linewidth=1.5) + geom_line(aes(x=-log(-log(U)),y=(z_p_0)),linewidth=1.5) + geom_line(aes(x=-log(-log(U)),y=(z_p_pos)),color="#c11432",linewidth=1.5) + 
  xlab(TeX("$-\\log(y_p)$")) +
  ylab(TeX("$z_p$" )) + theme_minimal() +
  annotate("text", x = 8, y = 0, label = TeX("$\\xi=-.2$")) +
  annotate("text", x = 8, y = 10, label = TeX("$\\xi=0$")) +
  annotate("text", x = 8, y = 25, label = TeX("$\\xi=.2$")) +
  theme_bw() +
  theme(panel.spacing = unit(2, "lines"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        strip.background = element_blank(),
        panel.border = element_rect(colour = "black", fill = NA))
```

### Maximum likelihood fitting of the GEV distribution

Try fitting GEV from the `ismev` package.

```{r}
data(venice)
gev.fit(venice[,2])
```

```{r}
df <- data.frame(x=1:(365*nrow(venice)), y=rnorm(n = 365*nrow(venice),mean = 0,sd = 1))
df$block_max <- rep(FALSE,365*nrow(venice))
for (i in 1:nrow(venice)) {
  df$block_max[df$y==(max(df$y[(365*(i-1)+1):(365*i)]))] <- TRUE
}
ggplot(df) + geom_point(aes(x=x,y=y,color=block_max)) + scale_color_manual(values=c("black","#C11432"))
```

```{r}

fit <- gev.fit(df$y[df$block_max==TRUE])
fit$mle

#gamGPDfit(x=df,datvar='y',threshold = 2.5,xiFrhs = xiFrhs)
```

Compare the fit with the density.

```{r}
g <- function(x) {
  1/fit$mle[2] * (1 + fit$mle[3]*((x - fit$mle[1])/fit$mle[2] )   )^(-(1+1/fit$mle[3]))* exp(-(1 + fit$mle[3]*((x - fit$mle[1])/fit$mle[2] )   )^(-1/fit$mle[3]) )
}
x_fit <- seq(1,4,0.01)
y_fit <- c()
for (i in seq_along(x_fit)) {
  y_fit[i] <- g(x_fit[i])
}
#plot(x_fit,y_fit)
df_fit <- data.frame(x_fit,y_fit)
ggplot() + geom_density(df %>% filter(block_max==TRUE),mapping=aes(y)) + geom_line(df_fit,mapping = aes(x_fit,y_fit),color='#c11432') + theme_minimal()
```

Try same fit with Venice data.

```{r}
venice <- data.frame(x=seq_along(venice[,2]),y=venice[,2])
fit <- gev.fit(venice[,2])
fit$mle
x_fit <- seq(70,200,1)
y_fit <- c()
g <- function(x) {
  1/fit$mle[2] * (1 + fit$mle[3]*((x - fit$mle[1])/fit$mle[2] )   )^(-(1+1/fit$mle[3]))* exp(-(1 + fit$mle[3]*((x - fit$mle[1])/fit$mle[2] ))^(-1/fit$mle[3]) )
}
for (i in seq_along(x_fit)) {
  y_fit[i] <- g(x_fit[i])
}
df_fit <- data.frame(x_fit,y_fit)
ggplot() + geom_density(venice,mapping=aes(y)) + geom_line(df_fit,mapping = aes(x_fit,y_fit),color='#c11432') + theme_minimal()

```

```{r}
N <- 1000
norm <- rnorm(N)
ex <- rexp(N)

plot(norm,ex)

hist(norm)
hist(ex)

u_1 <- pnorm(norm)
u_2 <- pexp(ex)

u_1 %>% head()
plot(u_1,u_2)

n_1 <- qnorm(u_1)
n_2 <- qnorm(u_2)
plot(n_1,n_2)

par_1 <- 1/(1-u_1)
par_2 <- 1/(1-u_2)
plot(par_1,par_2)
```

## Plot Gumbel distribution using PIT

```{r}
mu <- 1
sig <- 1
U <- runif(1000)
z <- mu - sig*log(-log(U))
df <- data.frame(z,U)
ggplot(df) + geom_point(aes(x=z,y=U))
```

## Try Port Pirie sea-levels diagnostic plots

Port Pirie data not available in the package, maybe try with Venice sea-levels data?

```{r}
data(portpirie)
venice[,2]
x=gev.fit(portpirie[,2])
gev.diag(x)
```

Try to recreate the likelihood function.

The function gets longer as initial values of the parameters are either pre-specified or calculated from the data. Therefore, `gev_fit.R` script contains the function.

```{r}
x <- portpirie[,2]
gev.diag(gev_fit(x))
# the function could return very high values 
# that cannot be minimum negative likelihood like 10^6
# would include in further development
```

Success! The plotting function works with the likelihood function and produces desired output.

### GEV plot functions

Now, try to replicate the four plots given above, starting with the probability plot. The model probability is a straight line, but empirical data will deviate from this line.

```{r}
# calculate GEV inverse CDF
GEV_inv <- function(U,mle) {
  mu <- mle[1]
  sig <- mle[2]
  xi <- mle[3]
 z <-  (sig/xi) * ((-log(U))^(-xi)-1) + mu 
 return(z)
}
# calculate GEV CDF
GEV <- function(z,mle) {
  mu <- mle[1]
  sig <- mle[2]
  xi <- mle[3]
  G <- c()
  for (i in seq_along(z))
 G[i] <- exp(  -(1 + xi*((z[i]-mu)/sig) )^(-1/xi)  )
 return(G)
}
```

Next is QQ plot. QQ plot compares two probability distributions by plotting their quantiles. In this example, it plots two GEV distributions with different parameters: empirical calculated though a formula used for the initial parameters in the optimisation and the model parameters (MLE).

```{r}
GEV_QQ <- function(x) {
  
# empirical distribution estimates
    sig_init <- sqrt(6 * var(x))/pi
    mu_init <- mean(x) - 0.57722 * sig_init
    xi_init <- 0.1
est_par <- c(mu_init,sig_init,xi_init)
# model (MLE) estimates
model_par <- gev_fit(x)$mle
# generate z using PIT for each case
q <- seq(from=1/(length(x)+1),to=length(x)/(length(x)+1),length.out=length(x)) #quantile
model <- GEV_inv(U=q,mle=model_par) #z generated from the model
est <- GEV_inv(U=q,mle=est_par) #z generated from GEV empirical estimates
dat <- x %>% sort()
return(data.frame(q=q,Model=model,Empirical=est,Data=dat))
}
```

The plot shows that empirical estimates are good since it is almost linear but shape parameter is a bit off, hence the slight curve (is it really the shape parameter causing this or just sampling variance?).

## Plot for UKCP data

Yearly maximum temperature data from the UKCP model.
```{r}
# calculate yearly maxima over 81 years
uk_winter <- readRDS("data/uk_1999_2079_winter.RDS")
uk_spring <- readRDS("data/uk_1999_2079_spring.RDS")
uk_summer <- readRDS("data/uk_1999_2079_summer.RDS")
uk_autumn <- readRDS("data/uk_1999_2079_autumn.RDS")
``` 

```{r}
# combine to 1 and take only London
tmp <- c()
for (i in 1:81) {
  tmp[((i-1)*360+1):(i*360)] <- c((uk_winter[uk_winter$is_location=="london",((90*(i-1))+7):((90*i)+6)] %>% as_vector),
                                      (uk_spring[uk_spring$is_location=="london",((90*(i-1))+7):((90*i)+6)] %>% as_vector),
                                      (uk_summer[uk_summer$is_location=="london",((90*(i-1))+7):((90*i)+6)] %>% as_vector),
                                      (uk_autumn[uk_autumn$is_location=="london",((90*(i-1))+7):((90*i)+6)] %>% as_vector))
}
for (i in 1:81) {
  tmp[((i-1)*360+1):(i*360)] <- c((uk_winter[uk_winter$is_location=="birmingham",((90*(i-1))+7):((90*i)+6)] %>% as_vector),
                                      (uk_spring[uk_spring$is_location=="birmingham",((90*(i-1))+7):((90*i)+6)] %>% as_vector),
                                      (uk_summer[uk_summer$is_location=="birmingham",((90*(i-1))+7):((90*i)+6)] %>% as_vector),
                                      (uk_autumn[uk_autumn$is_location=="birmingham",((90*(i-1))+7):((90*i)+6)] %>% as_vector))
}
# calculate yearly maxima
x <- c()
for (i in 1:81) {
  x[i] <- max(tmp[(360*(i-1)+1):(i*360)])
}
I <- seq_along(X)
gradient <- summary(lm(x~I))$coefficients[2,1]

shift <- c()
for (i in seq_along(x)) {
  shift[i] <- x[i] - (i-1)*gradient 
}

plot(I,x)
plot(I,shift)

# try shifting all the data with the corresponding yearly coefficients
# calculate yearly maxima
y <- tmp
I <- seq_along(y)
gradient <- summary(lm(y~I))$coefficients[2,1]

shift_tmp <- c()
for (i in seq_along(y)) {
  shift_tmp[i] <- y[i] - (i-1)*gradient 
}

plot(I,y)
plot(I,shift_tmp)

```

```{r}
x <- tmp
mle <- gev_fit(x)$mle
temp <- data.frame(Temperature=sort(x))
# calculate upper and lower bounds for 95% CI to add to the plots
Ulow<-sapply(seq_along(x), function(i){qbeta(0.025, i, length(x)+1-i)})
Uup<-sapply(seq_along(x), function(i){qbeta(0.975, i, length(x)+1-i)})
Model <- temp %>% dplyr::select(Temperature) %>% mutate(Model=GEV(z=Temperature,mle=mle)) %>% 
  mutate(Ulow=Ulow) %>% mutate(Uup=Uup) 

Empirical <- temp %>% dplyr::select(Temperature) %>% arrange(Temperature) %>% mutate(Empirical=row_number()/(nrow(temp)+1)) 
df <- temp %>% left_join(Empirical,by="Temperature") %>% left_join(Model,by="Temperature")
df %>% head()
p1 <- df %>% ggplot()  + geom_segment(aes(x = 0, y = 0, xend = 1, yend = 1), color = "#C11432",size=1.2)+ geom_point(aes(x=Empirical,y=Model)) + geom_line(aes(x=Empirical,y=Uup),linetype="dashed", color="#C11432")  + geom_line(aes(x=Empirical,y=Ulow), linetype="dashed", color="#C11432")  + 
  geom_ribbon(aes(x=Empirical,ymin=Uup,ymax=Ulow), fill="#C11432", alpha=0.2)+
  ggtitle("Probability Plot") + xlab("Empirical") + ylab("Model")
df <- GEV_QQ(sort(x))%>% mutate(GEV_Ulow=qgev(p=Ulow,loc=mle[1],scale=mle[2],shape=mle[3])) %>%  mutate(Ulow=Ulow) %>% mutate(Uup=Uup)%>% mutate(GEV_Uup=qgev(p=Uup,loc=mle[1],scale=mle[2],shape=mle[3]))
p2 <- df %>% ggplot()  + 
  geom_segment(aes(x = min(x), y = min(x), xend = max(x), yend = max(x)), color = "#C11432",size=1.2,alpha=0.5)+
  geom_point(aes(y=Data,x=Model)) + geom_line(aes(x=Model,y=GEV_Uup),linetype="dashed", color="#C11432")  + geom_line(aes(x=Model,y=GEV_Ulow), linetype="dashed", color="#C11432")  +
  geom_ribbon(aes(x=Model,ymin=GEV_Uup,ymax=GEV_Ulow), fill="#C11432", alpha=0.2)+
  ggtitle("Quantile Plot ") + ylab("Empirical") + xlab("Model") 
grid.arrange(p1,p2,ncol=2)
```



The next plot is return level plot.

```{r}
df1 <- df
x <- temp[,1]
#fit <- gev_fit(x)
fit <- as.vector(top_bottom[1,])
f <- seq(0.999,0.01,-0.001)
z <- GEV_inv(U=p,mle=fit$mle)
df <- data.frame(p=f,z)
#f <- p
q <- gevq(fit$mle, 1 - f)
d <- t(gev.rl.gradient(a = fit$mle, p = 1 - f))
v <- apply(d, 1, q.form, m = fit$cov)
CI <- data.frame(u= q + 1.96 * sqrt(v),l=q - 1.96 * sqrt(v))
# add also 95% bootstrap interval
 boot <-   rgev(length(x),loc=fit$mle[1],scale=fit$mle[2],shape=fit$mle[3])
 fit_boot <- gev_fit(boot)$mle
 
#  fit_boot <- data.frame(mu=numeric(),sig=numeric(),xi=numeric())
#   for (i in 1:1000) {
#      boot <-   rgev(length(x),loc=fit$mle[1],scale=fit$mle[2],shape=fit$mle[3])
#  fit_boot[nrow(fit_boot) +1,] <- gev_fit(boot)$mle
#  
#   }
#  
# fb <-  fit_boot %>% arrange(mu) %>% slice(26:975) 
# top_bottom <- fb %>% slice(c(1,n()))
# probboot <- pgev(q=boot,loc=fit$mle[1],scale=fit$mle[2],shape=fit$mle[3])  
# boot <- boot[probboot> 0.025 & probboot <0.975]
# probboot <- probboot[probboot> 0.025 & probboot <0.975]
# bootstrap <- data.frame(boot,probboot)
ggplot() + 
  geom_line(aes(x=-1/log(df$p),y=df$z)) +
 # geom_point(aes(x=-1/log(bootstrap$probboot),y=bootstrap$boot))+
  geom_line(aes(x=-1/log(df$p),y=CI$u),col="#C11432",linetype="dashed") +
  geom_line(aes(x=-1/log(df$p),y=CI$l),col="#C11432", linetype="dashed") +
  geom_ribbon(aes(x=-1/log(df$p),ymin=CI$u,ymax=CI$l), fill="#C11432", alpha=0.2)+
  geom_point(aes(x=-1/log(df1$q),y=df1$Data)) + coord_trans(x="log") +
  scale_x_continuous( breaks=c(10,100,500,1000)) +
  xlab("Return level (years)") +
  ylab(TeX("$z_p$"))   
```
#### Try the same for GPD

```{r}
u <- 32
x <- sort(tmp[tmp>u])
mle <- gpd.fit(x,threshold=u)$mle
# CI for xi
# f <- gpd.fit(x,threshold=u)
# c(f$mle[2]-qnorm(p=0.975)*f$se[2],f$mle[2]+qnorm(p=0.975)*f$se[2])

temp <- data.frame(Temperature=x)
# calculate GPD inverse CDF
GPD_inv <- function(mle,threshold,p) {
  sig <- mle[1]
  xi <- mle[2]
 z <- threshold + (sig * (p^(-xi) - 1))/xi
 return(z)
}
# calculate GPD CDF
GPD <- function(z,mle,threshold) {
  sig <- mle[1]
  xi <- mle[2]
  G <- c()
  for (i in seq_along(z))
 G[i] <- 1 - (1 + (xi * (z[i] - threshold))/sig)^(-1/xi)
 return(G)
}

GPD_QQ <- function(x,mle,threshold) {
# model (MLE) estimates
model_par <- mle
u <- threshold
# generate z using PIT for each case
q <-1- seq(from=1/(length(x)+1),to=length(x)/(length(x)+1),length.out=length(x)) #quantile
model <- GPD_inv(mle=model_par,threshold=u,p=q) #z generated from the model
# calculate upper and lower bounds for 95% CI to add to the plots
Ulow<-sapply(seq_along(x), function(i){qbeta(0.025, i, length(x)+1-i)})
Uup<-sapply(seq_along(x), function(i){qbeta(0.975, i, length(x)+1-i)})
Model <- temp %>% select(Temperature) %>% mutate(Model=GEV(z=Temperature,mle=mle)) %>% 
  mutate(Ulow=Ulow) %>% mutate(Uup=Uup) 
#est <- GEV_inv(U=q,mle=est_par) #z generated from GEV empirical estimates
dat <- x %>% sort()
return(data.frame(q=q,Model=model,Data=dat,Ulow=Ulow,Uup=Uup))
}

#Model <- temp %>% select(Temperature) %>% mutate(Model=GPD(z=Temperature,mle=mle,threshold=u))
# calculate upper and lower bounds for 95% CI to add to the plots
Ulow<-sapply(seq_along(x), function(i){qbeta(0.025, i, length(x)+1-i)})
Uup<-sapply(seq_along(x), function(i){qbeta(0.975, i, length(x)+1-i)})
Model <- temp %>%  select(Temperature) %>% mutate(Model=GPD(z=Temperature,mle=mle,threshold=u)) %>% 
  mutate(Ulow=Ulow) %>% mutate(Uup=Uup) 
Empirical <- temp %>% select(Temperature) %>% arrange(Temperature) %>% mutate(Empirical=row_number()/(nrow(temp)+1)) 
df <- temp %>% left_join(Empirical,by="Temperature") %>% left_join(Model,by="Temperature")
df %>% head()
p1 <- df %>% ggplot()  + geom_segment(aes(x = 0, y = 0, xend = 1, yend = 1), color = "#C11432",size=1.5)+ geom_point(aes(x=Empirical,y=Model)) +
  geom_line(aes(x=Empirical,y=Uup),linetype="dashed", color="#C11432")  + geom_line(aes(x=Empirical,y=Ulow), linetype="dashed", color="#C11432")  + 
  geom_ribbon(aes(x=Empirical,ymin=Uup,ymax=Ulow), fill="#C11432", alpha=0.2)+
  ggtitle("Probability Plot") + xlab("Empirical") + ylab("Model")
df <- GPD_QQ(x=temp[,1],mle=mle,threshold=u) %>% mutate(GPD_Ulow=qgpd(p=Ulow,loc=u,scale=mle[1],shape=mle[2])) %>%  mutate(Ulow=Ulow) %>% mutate(Uup=Uup)%>% mutate(GPD_Uup=qgpd(p=Uup,loc=u,scale=mle[1],shape=mle[2]))
x <- temp[,1]
p2 <- df %>% ggplot()  + 
  geom_segment(aes(x = min(x), y = min(x), xend = max(x), yend = max(x)), color = "#C11432",size=1.5,alpha=0.5)+
  geom_point(aes(y=Data,x=Model)) +
  geom_line(aes(x=Model,y=GPD_Uup),linetype="dashed", color="#C11432")  + geom_line(aes(x=Model,y=GPD_Ulow), linetype="dashed", color="#C11432")  +
  geom_ribbon(aes(x=Model,ymin=GPD_Uup,ymax=GPD_Ulow), fill="#C11432", alpha=0.2) + ggtitle("Quantile Plot ") + ylab("Empirical") + xlab("Model") 
grid.arrange(p1,p2,ncol=2)

```

```{r}
df1 <- df
x <- shift_tmp
u <- 30
fit <- gpd.fit(x,threshold = u)
p <- seq(0.999,0.01,-0.001)
z <- GPD_inv(mle=fit$mle,threshold=u,p=p) 
df <- data.frame(p,z)
a <- c(fit$rate,fit$mle)
u <- fit$threshold
n <- fit$n
la <- fit$rate
mat <- fit$cov
npy <- 360
dat <- x
xdat <- shift_tmp
  eps <- 1e-06
    a1 <- a
    a2 <- a
    a3 <- a
    a1[1] <- a[1] + eps
    a2[2] <- a[2] + eps
    a3[3] <- a[3] + eps
    jj <- seq(-1, 3.75 + log10(npy), by = 0.1)
    m <- c(1/la, 10^jj)
    q <- gpdq2(a[2:3], u, la, m)
    d <- t(gpd.rl.gradient(a = a, m = m))
    mat <- matrix(c((la * (1 - la))/n, 0, 0, 0, mat[1, 1], mat[1, 
        2], 0, mat[2, 1], mat[2, 2]), ncol = 3)
    v <- apply(d, 1, q.form, m = mat)
    plot(m/npy, q, log = "x", type = "n", xlim = c(0.1, max(m)/npy), 
        ylim = c(u, max(xdat, q[q > u - 1] + 1.96 * sqrt(v)[q > 
            u - 1])), xlab = "Return period (years)", ylab = "Return level", 
        main = "Return Level Plot")
    lines(m[q > u - 1]/npy, q[q > u - 1])
    lines(m[q > u - 1]/npy, q[q > u - 1] + 1.96 * sqrt(v)[q > 
        u - 1], col = 4)
    lines(m[q > u - 1]/npy, q[q > u - 1] - 1.96 * sqrt(v)[q > 
        u - 1], col = 4)
    nl <- n - length(dat) + 1
    sdat <- sort(xdat)
    points((1/(1 - (1:n)/(n + 1))/npy)[sdat > u], sdat[sdat > 
        u])


df <- data.frame(p= m[q > u - 1]/npy,z=(q[q > u - 1]) )
CI <- data.frame(p= m[q > u - 1]/npy,
                 l= (q[q > u - 1] - 1.96 * sqrt(v)[q > 
        u - 1]),
                 u= q[q > u - 1] + 1.96 * sqrt(v)[q > 
        u - 1])
df1 <- data.frame(q=(1/(1 - (1:n)/(n + 1))/npy)[sdat>u],
                    Data=sdat[sdat > u])



ggplot() + 
      xlim( c(0.3,1000)) +
  geom_line(aes(x=df$p,y=df$z)) +
  geom_line(aes(x=df$p,y=CI$u),col="#C11432",linetype="dashed") +
  geom_line(aes(x=df$p,y=CI$l),col="#C11432", linetype="dashed") +
  geom_ribbon(aes(x=df$p,ymin=CI$l,ymax=CI$u), fill="#C11432", alpha=0.2)+
  geom_point(aes(x=df1$q,y=df1$Data)) + coord_trans(x="log") +

  scale_x_continuous( breaks=c(10,100,500,1000),limits=c(0.3,1000)) +
  xlab("Return level (years)") +

  ylab(TeX("$z_p$"))   
```

### Time dependence study


```{r}
plot_frechet <- function(N) {
  set.seed(1)
 U_1 <- runif(N)
X_1 <- -1/(log(U_1))
U_2 <- runif(N)
X_2 <- -1/(log(U_2))
ggplot(data.frame(X_1=X_1/N,X_2=X_2/N)) + geom_point(aes(x=X_1,y=X_2),alpha=0.3,size=2) +
  xlab(TeX(paste0("$X_1/",N,"$"))) +
  ylab(TeX(paste0("$X_2/",N,"$"))) 
}

# plot radial and angular components
plot_r_w_frechet <- function(N) {
  set.seed(1)
 U_1 <- runif(N)
X_1 <- -1/(log(U_1))
U_2 <- runif(N)
X_2 <- -1/(log(U_2))
R <- (X_1+X_2)/N
W <- (X_1/N)/(R)
ggplot(data.frame(R=R,W=W)) + geom_point(aes(x=W,y=R),alpha=0.3,size=2) +
  ylab(TeX(paste0("$(X_1+X_2)/",N,"$"))) +
  xlab(TeX(paste0("$X_1/(X_1+X_2)$"))) 
}


#p1 <- plot_frechet(N=50)
p1 <- plot_frechet(N=50000)
p2 <- plot_r_w_frechet(N=50000)
grid.arrange(p1,p2,ncol=2)

# plot radial and angular components
U_1 <- runif(N)
X_1 <- -1/(log(U_1))
U_2 <- runif(N)
X_2 <- -1/(log(U_2))
R <- X_1/N+X_2/N
W <- (X_1/N)/(R)
plot(R,W)
X_3 <- X_1[R>0.1]
W_1 <- W[R>0.01]
R_1 <- R[R>0.01]
plot(log(R_1),W_1)
```

### 5.2 Maxima of stationary sequences

Simulate series $X_i=\max (Y_{i-1}, Y_i)$ where $Y$ has cdf $F_Y(y)= \exp \left\{ -\frac{1}{(a+1)y} \right\}$ for different values of $a$.

```{r}
# set a
a <- 0
# generate from Y
U <- runif(50)
Y <- c()
for (i in 1:50) {
Y[i] <- -1/(log(U[i])*(a+1))  
}
X <- Y[1]
for (i in 2:50) {
  X[i] <- max(Y[i-1],Y[i])
}
df <- data.frame(X,Y,ite=1:50)
x <- ggplot(df) + geom_point(aes(x=ite, y=X))
y <- ggplot(df) + geom_point(aes(x=ite, y=Y))
grid.arrange(x,y,ncol=2)
```

```{r}
# set a
a <- 1/3
# generate from Y
U <- runif(50)
Y <- c()
for (i in 1:50) {
Y[i] <- -1/(log(U[i])*(a+1))  
}
X <- Y[1]
for (i in 2:50) {
  X[i] <- max(Y[i-1],Y[i])
}
df <- data.frame(X,Y,ite=1:50)
x <- ggplot(df) + geom_point(aes(x=ite, y=X))
y <- ggplot(df) + geom_point(aes(x=ite, y=Y))
grid.arrange(x,y,ncol=2)
```

```{r}
# set a
a <- 2/3
# generate from Y
U <- runif(50)
Y <- c()
for (i in 1:50) {
Y[i] <- -1/(log(U[i])*(a+1))  
}
X <- Y[1]
for (i in 2:50) {
  X[i] <- max(Y[i-1],Y[i])
}
df <- data.frame(X,Y,ite=1:50)
x <- ggplot(df) + geom_point(aes(x=ite, y=X))
y <- ggplot(df) + geom_point(aes(x=ite, y=Y))
grid.arrange(x,y,ncol=2)
```

```{r}
# set a
a <- 1
# generate from Y
U <- runif(50)
Y <- c()
for (i in 1:50) {
Y[i] <- -1/(log(U[i])*(a+1))  
}
X <- Y[1]
for (i in 2:50) {
  X[i] <- max(Y[i-1],Y[i])
}
df <- data.frame(X,Y,ite=1:50)
x <- ggplot(df) + geom_point(aes(x=ite, y=X))
y <- ggplot(df) + geom_point(aes(x=ite, y=Y))
grid.arrange(x,y,ncol=2)
ggplot(df) + geom_point(aes(x=X, y=Y))
```

This makes sense. For example, $z=1$ corresponds to `r exp(-1/(1))` for $a=0$ and `r exp(-1/2)` for $a=1$. Because the quantile is lower for $a=0$, the distribution has more values larger then $1$ than the case when $a=1$.

### Extend the idea to a series of coefficients

Define $X_1=\max \left\{ a_1Y_1,a_2Y_2,\dots, a_dY_d \right\}$ and $X_2=\max \left\{ b_1Y_1,b_2Y_2,\dots, b_dY_d \right\}$. Further assume that $Y_i$ are independent with Fréchet distribution. Varying parameters are length $d$ taking max over, sequences $a$ and $b$.

```{r}
d=5
# generate Y
Y <- c()
for (i in 1:50) {
Y[i] <- -1/(log(U[i])*(a+1))  
}
X[1:(d-1)] <- Y[1:(d-1)]
for (i in d:50) {
  X[i] <- max(Y[(i-(d-1)):i])
}

df <- data.frame(X,Y,ite=1:50)
x <- ggplot(df) + geom_point(aes(x=ite, y=X))
y <- ggplot(df) + geom_point(aes(x=ite, y=Y))
grid.arrange(x,y,ncol=2)
# define a_1,...,a_50
#a_seq <- seq(1,0.1,length.out=50)
# define b_1,...,b_50
#b_seq <- seq(1,0.5,length.out=50)
ggplot(df) + geom_point(aes(x=X, y=Y))
```
Try different values of $d$.
```{r}
generate_dependent_X_Y <- function(N) {
  set.seed(1)
d <- 3
U <- runif(d*N)
a <- rep(1/3,d)
b <- (1:3)/6
# generate Y
Y <- c()
Y <- -1/(log(U) ) 
# generate X
X_1 <- c()
X_2 <- c()
for (j in 1:N) {
X_1[j] <- max(Y[(d*(j-1)+1):(d*j)]*a)/N
X_2[j] <- max(Y[(d*(j-1)+1):(d*j)]*b)/N
}

# plot X_1/N against X_2/N
df1 <- data.frame(X_1=X_1,X_2=X_2,uni=seq(0.02,1,0.02),
             uni_3=  seq(0.02,5/6,length.out=50), 
             u_x=seq(0,0.1,length.out=50),
             u_y=0.1-seq(0,0.1,length.out=50),
             line_1=seq(0.02,1,0.02)*(1/2),line_2=seq(0.02,1,0.02)*(1),line_3=seq(0.02,5/6,length.out=50)*(3/2))
df <-df1 %>%  mutate(above_u=(X_1+X_2>0.1))  
return(df)
}
plot_dependence <- function(df) {
N <- nrow(df)
ggplot(df) + geom_point(aes(x=X_1, y=X_2,col=above_u),alpha=0.5) +
  scale_color_manual(values = c("TRUE" = "#C11432",
                                "FALSE"="black")) +
   labs(color=TeX('Above $u$')) +
   xlab(TeX(paste0("$X_1/",N,"$"))) +
  ylab(TeX(paste0("$X_2/",N,"$"))) +
  geom_line(aes(x=uni,y=line_1),alpha=0.3) +
    geom_line(aes(x=uni,y=line_2),alpha=0.3) +
    geom_line(aes(x=uni_3,y=line_3),alpha=0.3) +
 # ggtitle(TeX(paste0("$N=",N,"$")))+
   geom_line(aes(x=u_x,y=u_y),col="#C11432",alpha=0.6,linetype="dashed") +
  annotate("text",x=0.12,y=0.05,label="u",col="#C11432") +
  coord_fixed()
}

p1 <- generate_dependent_X_Y(N=50) %>%  plot_dependence() 
p2 <- generate_dependent_X_Y(N=50000) %>% plot_dependence() 
grid.arrange(p1,p2,ncol=2)
```
Plot only above threshold?

Plot radial and angular components.

```{r}
plot_dependence_R_W <- function(df) {
  df1 <- data.frame(r=(df$X_1+df$X_2),w=df$X_1/(df$X_1+df$X_2),above_u=df$above_u)%>% mutate(N=rep(paste0("n=",nrow(df)),nrow(df))) %>% 
  mutate(u_x=seq(0,1,length.out=nrow(df))) %>% 
  mutate(u_y=rep(0.1,nrow(df)))
  ggplot(df1) + geom_point(data=df1,aes(x=w,y=log(r),col=above_u),alpha=0.5)+
  geom_line(data=df1,aes(x=u_x,y=log(u_y)),col="#C11432",alpha=0.6,linetype="dashed") +
  annotate("segment",x=0.4,xend=0.4,y=-12.5,yend=1,alpha=0.3) +
    annotate("segment",x=1/2,xend=1/2,y=-12.5,yend=1,alpha=0.3) +
    annotate("segment",x=2/3,xend=2/3,y=-12.5,yend=1,alpha=0.3) +
  xlab(TeX("$W$")) +   ylab(TeX("$\\log (R)$")) +
    ylim(c(-12.5,1))+
  annotate("text",x=1,y=-2.7,label="u",col="#C11432") +
scale_color_manual(values = c("TRUE" = "#C11432",
                                "FALSE"="black")) +
   labs(color=TeX('Above $u$'))
}

p1 <- plot_dependence_R_W(df=generate_dependent_X_Y(N=50))
p2 <- plot_dependence_R_W(df=generate_dependent_X_Y(N=50000))
grid.arrange(p1,p2,ncol=2)
```

 
### Transformation to Gumbel margins

Use PIT to transform Port Pirie data to have Gumbel margins (as motivated by Heffernan and Tawn (2004)).

```{r}
# create x and y quantile variables
data(portpirie)
portpirie <- data.frame(SeaLevel=portpirie,Year=seq(1,length(portpirie),1))
U <- portpirie %>% select(SeaLevel,Year) %>% arrange(SeaLevel) %>% mutate(u=row_number()/(nrow(portpirie)+1))
# generate Gumbel margins

df <- portpirie %>% left_join((U %>% select(Year,u)),by="Year") 
df <- df %>% mutate(Gumbel=-log(-log(u)))
df %>% head()

ggplot(df)+ geom_point(aes(x=Year,y=SeaLevel)) + geom_point(aes(x=Year,y=u),col="#C11432") +
  geom_point(aes(x=Year,y=Gumbel),col="#009ADA") +
  theme(aspect.ratio = 1) + xlab("Year") + ylab("Maximum Sea Level")
```



