---
title: "Exploration of potential parametric forms of conditional extreme value models"
output: html_document
date: "2024-10-25"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(MASS)
library(tidyverse)
library(latex2exp)
library(viridis)
library(plgp)
library(gridExtra)
library(here)
library(tmap)
library(units)
library(sf)
```

## Set-up

Currently, we work with 19 years of daily maximum temperature data over summer months (June, July, August). (The issue with year 20 remains unresolved.)

```{r}

```

## Follow-up from meeting with Simon: exploring $\alpha$ against distance

Recall the plot of marginal estimates of $\alpha$ parameter againt distance[m] for the three conditioning sites.

```{r}
# pick a function of alpha as a divide: diff for each site 

```

## Sequential parameter estimation

1. Fix $\beta=0$, estimate $\hat{\alpha}$.
2. Fix $\alpha=\hat{\alpha}$, estimate $\hat{\beta}$.
3a. Fix $\alpha=\hat{\alpha}$ and $\beta=\hat{\beta}$, estimate $\hat{\mu}$ and $\hat{\sigma}$.
3b. Calculate observed residuals, estimate $\hat{\mu_{AGG}}$, $\hat{\sigma_{AGG}}$, $\hat{\delta}_l$ and $\hat{\delta}_u$.


```{r}

```

## Conditioning on any random site

Currently, the analysis is conditioning on Birmingham, Glasgow and London, ordered as first 3 columns of the temperature data for an easy link. 