---
title: "Vine_copula_explore"
output: html_document
date: "2024-10-22"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# libraries and functions ----
library(VineCopula)
library(rvinecopulib)
library(evd)
library(ismev)
library(tidyverse)
library(MASS) # use dplyr::select to avoid function conflict
library(pracma) # incomplete gamma function, will be used later in cond.ext.model for modelling the residual margins
library(texmex) # air pollution data
file.sources = list.files(pattern="*helpers.R")
sapply(file.sources,source,.GlobalEnv)

# set theme defaults to be black rectangle
theme_set(theme_bw())
theme_replace(
  panel.spacing = unit(2, "lines"),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  strip.background = element_blank(),
  panel.border = element_rect(colour = "black", fill = NA) )
```

## Fit vine copula to winter data

```{r}
fitw <- RVineStructureSelect(as.data.frame(winter %>% apply(c(2),FUN=row_number)/(nrow(winter)+1)),
                     trunclevel = 3, indeptest = FALSE)
fitw
```

Now fit summer pollution data vine copula parameters but impose the tree structure from the winter pollution data.

```{r}
fits1 <- RVineStructureSelect(as.data.frame(summer %>% apply(c(2),FUN=row_number)/(nrow(summer)+1)),
                     trunclevel = 3, indeptest = FALSE)
# with imposed tree structure
fits2 <- RVineCopSelect(data = as.data.frame(summer %>% apply(c(2),FUN=row_number)/(nrow(summer)+1)), Matrix = fitw$Matrix,
                     trunclevel = 3, indeptest = FALSE)

```
First, the model as previously done.
```{r}
fits1
```
Second, our new model with imposed tree structure.

```{r}
fits2
```

Compare the two fits using the Vuong test.

```{r}
Vuong <- RVineVuongTest(data= as.data.frame(summer %>% apply(c(2),FUN=row_number)/(nrow(summer)+1)),RVM1 = fits1, RVM2 = fits2)
data.frame("statistic"=Vuong[[1]],"p.value"=Vuong[[4]])
```

$p$-value is very small, hence we reject the NULL hypothesis. The statistic value is positive, hence we prefer RVM1 without the imposed tree structure.

Do the same the other way around, i.e., start by fitting vine copula to summer pollution data.

```{r}
fits1
```

```{r}
# with imposed tree structure
fitw2 <- RVineCopSelect(data = as.data.frame(winter %>% apply(c(2),FUN=row_number)/(nrow(winter)+1)), Matrix = fits1$Matrix,
                     trunclevel = 3, indeptest = FALSE)
fitw # no imposed structure
fitw2 # imposed tree structure
```
 Compare the two fits for winter pollution data.
 
```{r}
Vuong <- RVineVuongTest(data= as.data.frame(winter %>% apply(c(2),FUN=row_number)/(nrow(winter)+1)),RVM1 = fitw, RVM2 = fitw2)
data.frame("statistic"=Vuong[[1]],"p.value"=Vuong[[4]])
```

In this case, $p$-value is high and $\nu$ statistic is close to zero, hence there is no difference between the model. Therefore, we conclude that we can impose summer tree structure to winter data but not the other way.

# Try combining the data on Uniform margins

```{r}
sw <- rbind(as.data.frame(summer %>% apply(c(2),FUN=row_number)/(nrow(summer)+1)),as.data.frame(winter %>% apply(c(2),FUN=row_number)/(nrow(winter)+1)))

fitsw <- RVineStructureSelect(sw,trunclevel = 3, indeptest = FALSE)
fitsw
```


## Fitting a vine copula to the observed residuals

`obsress` ia a matrix of observed residuals on the uniform margins for the summer pollution dataset conditioning on $O_3$ being extreme.

```{r}
v <- 0.7
# transform to Laplace margins
winter_lap <- as.data.frame((winter %>% apply(c(2),FUN=row_number))/(nrow(winter)+1)) %>% apply(c(1,2),FUN=unif_laplace_pit) %>% as.data.frame()
summer_lap <- as.data.frame((summer %>% apply(c(2),FUN=row_number))/(nrow(summer)+1)) %>% apply(c(1,2),FUN=unif_laplace_pit) %>% as.data.frame()

obsresw <- (observed_residuals(df = winter_lap,given = 1,v = v) %>% apply(c(2),FUN=row_number))/(nrow(winter_lap)*(1-v)+1)

obsress <- (observed_residuals(df = summer_lap,given = 1,v = v) %>% apply(c(2),FUN=row_number))/(nrow(summer_lap)*(1-v)+1)

fitrw <- RVineStructureSelect(obsresw,
                     trunclevel = 3, indeptest = TRUE)
fitrs <- RVineStructureSelect(obsress,
                     trunclevel = 3, indeptest = TRUE)
# with imposed tree structure
fitrs2 <- RVineCopSelect(data = obsress, Matrix = fitrw$Matrix,
                     trunclevel = 3, indeptest = FALSE)
fitrw2 <- RVineCopSelect(data = obsresw, Matrix = fitrs$Matrix,
                     trunclevel = 3, indeptest = FALSE)
fitrw1 <- rvinecopulib::vinecop(data=obsresw)
fitrw1
```

 Summer observed residual vine copula structure.
 
```{r}
fitrs
```
 
Summer observed residuals vine cop parameters with imposed winter copula structure.
 
```{r}
fitrs2
```
 
Winter observed residual vine copula structure.
  
```{r}
fitrw
```

Winter observed residuals vine cop parameters with imposed summer copula structure.
 
```{r}
fitrw2
```

### Try Vuong test on observed residuals vine copulas.

Summer: no significant difference, can impose winter tree structure.
```{r}
Vuong <- RVineVuongTest(data= obsress,RVM1 = fitrs, RVM2 = fitrs2)
data.frame("statistic"=Vuong[[1]],"p.value"=Vuong[[4]])
```
and winter: again no significant difference.

```{r}
Vuong <- RVineVuongTest(data= obsresw,RVM1 = fitrw, RVM2 = fitrw2)
data.frame("statistic"=Vuong[[1]],"p.value"=Vuong[[4]])
```

Vuong test compares whether model $2$ is an improvement to model $1$. Since p values are not below significance level $0.05$, we accept a model with imposed structure?

## Model comparison without normality assumption

We run this for the observed residuals of the winter pollution dataset conditioning on $O_3$ above $0.7$ quantile threshold.

```{r}
# simulate from a vine copula the ranks of the data
N_sim <- dim(obsresw)[1]

# find the simulated values for each of the models
Zsim1 <- RVineSim(N=N_sim,RVM=fitrw)
Zsim2 <- RVineSim(N=N_sim,RVM=fitrw2)
Vuongstat <- function(Zsim1,Zsim2,Nrep) {
Zsim <- rbind(Zsim1,Zsim2)
statistic <- c()
for (i in 1:Nrep) {
  # randomly sample with replacement
  row_sample <- sample(1:(2*N_sim), size=N_sim,replace=FALSE)
  Y1 <- Zsim[row_sample,]
  Y2 <- Zsim[-row_sample,]
  # fit model for each
  fit1 <- RVineStructureSelect(Y1,trunclevel = 3, indeptest = FALSE)
  fit2 <- RVineStructureSelect(Y2,trunclevel = 3, indeptest = FALSE)
  # calculate Vuong test statistic
  statistic[i] <- RVineVuongTest(data=obsresw,RVM1 = fit1,RVM2 = fit2)$statistic
}
}
Vuong_statistic <- statistic
hist(Vuong_statistic,n=20)
abline(v=Vuong[[1]], col="#C11432",lty= "dashed")
```


```{r, echo= FALSE}
# write into a function for conditioning on other pollutants

fn <- function(j) {
obsresw <- (observed_residuals(df = winter_lap,given = j,v = v) %>% apply(c(2),FUN=row_number))/(nrow(winter_lap)*(1-v)+1)

obsress <- (observed_residuals(df = summer_lap,given = 1,v = v) %>% apply(c(2),FUN=row_number))/(nrow(summer_lap)*(1-v)+1)

fitrw <- RVineStructureSelect(obsresw,
                     trunclevel = 3, indeptest = TRUE)
fitrs <- RVineStructureSelect(obsress,
                     trunclevel = 3, indeptest = TRUE)
# with imposed tree structure
fitrs2 <- RVineCopSelect(data = obsress, Matrix = fitrw$Matrix,
                     trunclevel = 3, indeptest = FALSE)
fitrw2 <- RVineCopSelect(data = obsresw, Matrix = fitrs$Matrix,
                     trunclevel = 3, indeptest = FALSE)
testw <- RVineVuongTest(data= obsresw,RVM1 = fitrw, RVM2 = fitrw2)
tests <- RVineVuongTest(data= obsress,RVM1 = fitrs, RVM2 = fitrs2)
return(list(fitrw,fitrw2,fitrs,fitrs2,testw,tests))
}
```

Conditioning on $NO_2$.

```{r}
x <- fn(j=2)
x
```

Conditioning on $NO$.

```{r}
x <- fn(j=3)
x
```

Conditioning on $SO_2$.

```{r}
x <- fn(j=4)
x
```

Conditioning on $PM_{10}$.

```{r}
x <- fn(j=5)
x
```

# Combining summer and winter data

```{r}
sw1 <- sw
fnsw <- function(j) {
  sw <- as.data.frame((sw %>% apply(c(2),FUN=row_number))/(nrow(sw)+1)) %>% apply(c(1,2),FUN=unif_laplace_pit) %>% as.data.frame()
obsresw <- (observed_residuals(df = sw,given = j,v = v) %>% apply(c(2),FUN=row_number))/(nrow(sw)*(1-v)+1)

obsress <- (observed_residuals(df = sw,given = 1,v = v) %>% apply(c(2),FUN=row_number))/(nrow(sw)*(1-v)+1)

fitrw <- RVineStructureSelect(obsresw,
                     trunclevel = 3, indeptest = TRUE)
fitrs <- RVineStructureSelect(obsress,
                     trunclevel = 3, indeptest = TRUE)

return(list(fitrw,fitrs))
}
fnsw(j=1)
```
We can perform the above comparison again for the observed residuals of joint summer and winter data conditional on each of the pollutants.