---
title: "Vine_copula_explore"
output: html_document
date: "2024-10-22"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# libraries and functions ----
library(VineCopula)
library(evd)
library(ismev)
library(tidyverse)
library(MASS) # use dplyr::select to avoid function conflict
library(pracma) # incomplete gamma function, will be used later in cond.ext.model for modelling the residual margins
library(texmex) # air pollution data
file.sources = list.files(pattern="*helpers.R")
sapply(file.sources,source,.GlobalEnv)

# set theme defaults to be black rectangle
theme_set(theme_bw())
theme_replace(
  panel.spacing = unit(2, "lines"),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  strip.background = element_blank(),
  panel.border = element_rect(colour = "black", fill = NA) )
```

## Fit vine copula to winter data

```{r}
fitw <- RVineStructureSelect(as.data.frame(winter %>% apply(c(2),FUN=row_number)/(nrow(winter)+1)),
                     trunclevel = 3, indeptest = FALSE)
fitw
```

Now fit summer pollution data vine copula parameters but impose the tree structure from the winter pollution data.

```{r}
fits1 <- RVineStructureSelect(as.data.frame(summer %>% apply(c(2),FUN=row_number)/(nrow(summer)+1)),
                     trunclevel = 3, indeptest = FALSE)
# with imposed tree structure
fits2 <- RVineCopSelect(data = as.data.frame(summer %>% apply(c(2),FUN=row_number)/(nrow(summer)+1)), Matrix = fitw$Matrix,
                     trunclevel = 3, indeptest = FALSE)

```
First, the model as previously done.
```{r}
fits1
```
Second, our new model with imposed tree structure.

```{r}
fits2
```

Compare the two fits using the Vuong test.

```{r}
RVineVuongTest(data= as.data.frame(summer %>% apply(c(2),FUN=row_number)/(nrow(summer)+1)),RVM1 = fits1, RVM2 = fits2)
```

Do the same the other way around, i.e., start by fitting vine copula to summer pollution data.

```{r}
fits1
```

```{r}
# with imposed tree structure
fitw2 <- RVineCopSelect(data = as.data.frame(winter %>% apply(c(2),FUN=row_number)/(nrow(winter)+1)), Matrix = fits1$Matrix,
                     trunclevel = 3, indeptest = FALSE)
fitw
fitw2
```
 Compare the two fits for winter pollution data.
```{r}
RVineVuongTest(data= as.data.frame(winter %>% apply(c(2),FUN=row_number)/(nrow(winter)+1)),RVM1 = fitw, RVM2 = fitw2)
```
 
 