---
title: "Vine_copula_explore"
output: html_document
date: "2024-10-22"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# libraries and functions ----
library(VineCopula)
library(evd)
library(ismev)
library(tidyverse)
library(MASS) # use dplyr::select to avoid function conflict
library(pracma) # incomplete gamma function, will be used later in cond.ext.model for modelling the residual margins
library(texmex) # air pollution data
file.sources = list.files(pattern="*helpers.R")
sapply(file.sources,source,.GlobalEnv)

# set theme defaults to be black rectangle
theme_set(theme_bw())
theme_replace(
  panel.spacing = unit(2, "lines"),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  strip.background = element_blank(),
  panel.border = element_rect(colour = "black", fill = NA) )
```

## Fit vine copula to winter data

```{r}
fitw <- RVineStructureSelect(as.data.frame(winter %>% apply(c(2),FUN=row_number)/(nrow(winter)+1)),
                     trunclevel = 3, indeptest = FALSE)
fitw
```

Now fit summer pollution data vine copula parameters but impose the tree structure from the winter pollution data.

```{r}
fits1 <- RVineStructureSelect(as.data.frame(summer %>% apply(c(2),FUN=row_number)/(nrow(summer)+1)),
                     trunclevel = 3, indeptest = FALSE)
# with imposed tree structure
fits2 <- RVineCopSelect(data = as.data.frame(summer %>% apply(c(2),FUN=row_number)/(nrow(summer)+1)), Matrix = fitw$Matrix,
                     trunclevel = 3, indeptest = FALSE)

```
First, the model as previously done.
```{r}
fits1
```
Second, our new model with imposed tree structure.

```{r}
fits2
```

Compare the two fits using the Vuong test.

```{r}
RVineVuongTest(data= as.data.frame(summer %>% apply(c(2),FUN=row_number)/(nrow(summer)+1)),RVM1 = fits1, RVM2 = fits2)
```

Do the same the other way around, i.e., start by fitting vine copula to summer pollution data.

```{r}
fits1
```

```{r}
# with imposed tree structure
fitw2 <- RVineCopSelect(data = as.data.frame(winter %>% apply(c(2),FUN=row_number)/(nrow(winter)+1)), Matrix = fits1$Matrix,
                     trunclevel = 3, indeptest = FALSE)
fitw
fitw2
```
 Compare the two fits for winter pollution data.
```{r}
RVineVuongTest(data= as.data.frame(winter %>% apply(c(2),FUN=row_number)/(nrow(winter)+1)),RVM1 = fitw, RVM2 = fitw2)
```
# Try combining the data on Uniform margins
```{r}
sw <- rbind(as.data.frame(summer %>% apply(c(2),FUN=row_number)/(nrow(summer)+1)),as.data.frame(winter %>% apply(c(2),FUN=row_number)/(nrow(winter)+1)))

fitsw <- RVineStructureSelect(sw,trunclevel = 3, indeptest = FALSE)
fitsw
```


## Fitting a vine copula to the observed residuals

`fitress` is a matrix of observed residuals on the uniform margins for the summer pollution dataset conditioning on $O_3$ being extreme.

```{r}
v <- 0.7
# transform to Laplace margins
winter_lap <- as.data.frame((winter %>% apply(c(2),FUN=row_number))/(nrow(winter)+1)) %>% apply(c(1,2),FUN=unif_laplace_pit) %>% as.data.frame()
summer_lap <- as.data.frame((summer %>% apply(c(2),FUN=row_number))/(nrow(summer)+1)) %>% apply(c(1,2),FUN=unif_laplace_pit) %>% as.data.frame()

obsresw <- (observed_residuals(df = winter_lap,given = 1,v = v) %>% apply(c(2),FUN=row_number))/(nrow(winter_lap)*(1-v)+1)

obsress <- (observed_residuals(df = summer_lap,given = 1,v = v) %>% apply(c(2),FUN=row_number))/(nrow(summer_lap)*(1-v)+1)

fitrw <- RVineStructureSelect(obsresw,
                     trunclevel = 3, indeptest = TRUE)
fitrs <- RVineStructureSelect(obsress,
                     trunclevel = 3, indeptest = TRUE)
# with imposed tree structure
fitrs2 <- RVineCopSelect(data = obsress, Matrix = fitrw$Matrix,
                     trunclevel = 3, indeptest = FALSE)
fitrw2 <- RVineCopSelect(data = obsresw, Matrix = fitrs$Matrix,
                     trunclevel = 3, indeptest = FALSE)
```

 Summer observed residual vine copula structure.
 
```{r}
fitrs
```
 
Summer observed residuals vine cop parameters with imposed winter copula structure.
 
```{r}
fitrs2
```
 
Winter observed residual vine copula structure.
  
```{r}
fitrw
```

Winter observed residuals vine cop parameters with imposed summer copula structure.
 
```{r}
fitrw2
```

### Try Vuong test on observed residuals vine copulas.

Summer
```{r}
RVineVuongTest(data= obsress,RVM1 = fitrs, RVM2 = fitrs2)
```
and winter

```{r}
RVineVuongTest(data= obsresw,RVM1 = fitrw, RVM2 = fitrw2)
```

Vuong test compares whether model $2$ is an improvement to model $1$. Since p values are not below significance level $0.05$, we accept a model with imposed structure?

```{r, echo= FALSE}
# write into a function for conditioning on other pollutants

fn <- function(j) {
obsresw <- (observed_residuals(df = winter_lap,given = j,v = v) %>% apply(c(2),FUN=row_number))/(nrow(winter_lap)*(1-v)+1)

obsress <- (observed_residuals(df = summer_lap,given = 1,v = v) %>% apply(c(2),FUN=row_number))/(nrow(summer_lap)*(1-v)+1)

fitrw <- RVineStructureSelect(obsresw,
                     trunclevel = 3, indeptest = TRUE)
fitrs <- RVineStructureSelect(obsress,
                     trunclevel = 3, indeptest = TRUE)
# with imposed tree structure
fitrs2 <- RVineCopSelect(data = obsress, Matrix = fitrw$Matrix,
                     trunclevel = 3, indeptest = FALSE)
fitrw2 <- RVineCopSelect(data = obsresw, Matrix = fitrs$Matrix,
                     trunclevel = 3, indeptest = FALSE)
testw <- RVineVuongTest(data= obsresw,RVM1 = fitrw, RVM2 = fitrw2)
tests <- RVineVuongTest(data= obsress,RVM1 = fitrs, RVM2 = fitrs2)
return(list(fitrw,fitrw2,fitrs,fitrs2,testw,tests))
}
```

Conditioning on $NO_2$.

```{r}
x <- fn(j=2)
x
```

Conditioning on $NO$.

```{r}
x <- fn(j=3)
x
```

Conditioning on $SO_2$.

```{r}
x <- fn(j=4)
x
```

Conditioning on $PM_{10}$.

```{r}
x <- fn(j=5)
x
```

```{r}
sw1 <- sw
fnsw <- function(j) {
  sw <- as.data.frame((sw %>% apply(c(2),FUN=row_number))/(nrow(sw)+1)) %>% apply(c(1,2),FUN=unif_laplace_pit) %>% as.data.frame()
obsresw <- (observed_residuals(df = sw,given = j,v = v) %>% apply(c(2),FUN=row_number))/(nrow(sw)*(1-v)+1)

obsress <- (observed_residuals(df = sw,given = 1,v = v) %>% apply(c(2),FUN=row_number))/(nrow(sw)*(1-v)+1)

fitrw <- RVineStructureSelect(obsresw,
                     trunclevel = 3, indeptest = TRUE)
fitrs <- RVineStructureSelect(obsress,
                     trunclevel = 3, indeptest = TRUE)

return(list(fitrw,fitrs))
}
fnsw(j=1)
```
