---
title: "bivariate_model_example"
output: html_document
date: "2025-03-03"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(pracma) # incomplete gamma function, will be used later in cond.ext.model for modelling the residual margins
library(here)
# set theme defaults to be black rectangle
theme_set(theme_bw())
theme_replace(
  panel.spacing = unit(2, "lines"),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  strip.background = element_blank(),
  panel.border = element_rect(colour = "black", fill = NA) )

source("../sample_distribution_helpers.R")
source("../likelihood_helpers.R")
source("../cond_model_helpers.R")
```

## Bivariate example of conditional extreme value model

This example uses a sample from a bivariate logistic extreme value distribution with Fréchet margins (then transformed to Laplace for modelling).

```{r}
set.seed(11)
N <- 5000
v <- 0.9 # threshold for conditioning variable
sims <- generate_Y(N=N) %>% link_log(dep=1/2) %>%
   apply(c(1,2),FUN=frechet_laplace_pit) %>% # transform to Laplace margins
  as.data.frame()
as_tibble(sims)
```

## Estimating parameters

`margin="AGG"` means Normal distribution is used for the residual margin (estimates are columns `mu_agg`, `sig_l`, `sig_u`, `delta_l` and `delta_u`).

`method=sequential2` mean that we fix $\beta=0$ and estimate $\hat{\alpha}$ (column `a` and `lika` for NLL (negative log-likelihood)), then fix $\alpha=\hat{\alpha}$ and estimate $\hat{\beta}$ (column `b` and `likb` for LL).

`keef_constraints=c(1,2)` means we use both Keef constraints as upper bound on $\beta$, i.e. $\beta_{max} = \min \left\{ \beta_{constraint1},\beta_{constraint2}  \right\}$.

`par_est()` function is from `cond_model_helpers.R` (a bit messy atm).

```{r}
par_est(df=sims,v=v,given=c(1),margin = "Normal", method = "sequential2", keef_constraints = c(1,2))
```


