---
title: "Explore rvinecopulib"
output: html_document
date: "2025-03-10"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# libraries and functions ----
library(VineCopula)
library(rvinecopulib)
#library(evd)
#library(ismev)
library(tidyverse)
#library(MASS) # use dplyr::select to avoid function conflict
library(pracma) # incomplete gamma function, will be used later in cond.ext.model for modelling the residual margins
library(texmex) # air pollution data
source("../sample_distribution_helpers.R")
source("../likelihood_helpers.R")
source("../cond_model_helpers.R")

# set theme defaults to be black rectangle
theme_set(theme_bw())
theme_replace(
  panel.spacing = unit(2, "lines"),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  strip.background = element_blank(),
  panel.border = element_rect(colour = "black", fill = NA) )
```

## Exploring `rvinecopulib` package

```{r}
v <- 0.7
# transform to Laplace margins
winter_lap <- as.data.frame((winter %>% apply(c(2),FUN=row_number))/(nrow(winter)+1)) %>% apply(c(1,2),FUN=unif_laplace_pit) %>% as.data.frame()
summer_lap <- as.data.frame((summer %>% apply(c(2),FUN=row_number))/(nrow(summer)+1)) %>% apply(c(1,2),FUN=unif_laplace_pit) %>% as.data.frame()
j <- 1
pew <-  par_est(df=winter_lap,v=v,given=j,margin = "AGG", method = "sequential2")
obsresw <- (observed_residuals(df = winter_lap,given = j,v = v,a = pew$a,b=pew$b) %>% apply(c(2),FUN=row_number))/(nrow(winter_lap)*(1-v)+1)

pes <-  par_est(df=summer_lap,v=v,given=j,margin = "AGG", method = "sequential2")
obsress <- (observed_residuals(df = summer_lap,given = j,v = v,a = pes$a,b=pes$b) %>% apply(c(2),FUN=row_number))/(nrow(summer_lap)*(1-v)+1)

```

## Estimate copula conditional on $O_3$

```{r}
vc1 <- rvinecopulib::vinecop(data=obsresw)
vc1
summary(vc1)
# compare with VineCopula fit
vc2 <- VineCopula::RVineStructureSelect(data=obsresw)
summary(vc2)
```

Compare also with using the independence test.

```{r}
vc1 <- rvinecopulib::vinecop(data=obsresw)
vc1
summary(vc1)
# compare with VineCopula fit
vc2 <- VineCopula::RVineStructureSelect(data=obsresw, indeptest = TRUE)
summary(vc2)
```

Winter pollution data, cond on $O_3$ with imposed summer pollution observed residuals tree structure.

```{r}
vcs1 <- rvinecopulib::vinecop(data=obsress)
vcs2 <- VineCopula::RVineStructureSelect(data=obsress)
vcw1 <- rvinecopulib::vinecop(data=obsresw,structure=get_structure(vcs1) )
vcw2 <- VineCopula::RVineCopSelect(data = obsresw,Matrix = vcs2$Matrix)
vcw1
summary(vcw1)
summary(vcw2)
```

## Copula statistic comparison

Compare negative log-likelihood, AIC and BIC for each of the copulas for both winter and summer data.There are more observations in summer, hence NLL values will be likely higher.

```{r}
tmp <- data.frame(cond_pollutant=character(),method=character(),season=character(),NLL=numeric(),AIC=numeric(),BIC=numeric())

AIC <- function(vc) {
  2*vc$npars+2*vc$loglik
}

BIC <- function(vc) {
  vc$npars*log(vc$nobs) +2*vc$loglik
}

for (i in 1:5) {
pew <-  par_est(df=winter_lap,v=v,given=i,margin = "AGG", method = "sequential2")
obsresw <- (observed_residuals(df = winter_lap,given = i,v = v,a = pew$a,b=pew$b) %>% apply(c(2),FUN=row_number))/(nrow(winter_lap)*(1-v)+1)

pes <-  par_est(df=summer_lap,v=v,given=i,margin = "AGG", method = "sequential2")
obsress <- (observed_residuals(df = summer_lap,given = i,v = v,a = pes$a,b=pes$b) %>% apply(c(2),FUN=row_number))/(nrow(summer_lap)*(1-v)+1)

vcw1 <- rvinecopulib::vinecop(data=obsresw)
# compare with VineCopula fit
vcw2 <- VineCopula::RVineStructureSelect(data=obsresw)

vcs1 <- rvinecopulib::vinecop(data=obsress)
# compare with VineCopula fit
vcs2 <- VineCopula::RVineStructureSelect(data=obsress)

tmp <- rbind(tmp,data.frame(cond_pollutant=names(winter)[i],method="rvinecopulib",season="winter",NLL=vcw1$loglik,AIC=AIC(vc=vcw1),BIC=BIC(vcw1)),
             data.frame(cond_pollutant=names(winter)[i],method="VineCopula",season="winter",NLL=vcw2$logLik,AIC=-vcw2$AIC,BIC=-vcw2$BIC),
             data.frame(cond_pollutant=names(winter)[i],method="rvinecopulib",season="summer",NLL=vcs1$loglik,AIC=AIC(vc=vcs1),BIC=BIC(vcs1)),
             data.frame(cond_pollutant=names(winter)[i],method="VineCopula",season="summer",NLL=vcs2$logLik,AIC=-vcs2$AIC,BIC=-vcs2$BIC)
             )
}

```

Plot these for comparison.

Compare likelihoods overall for each method and each season.

```{r}
tmp <- tmp %>% mutate(season=factor(season),method=factor(method))
ggplot(tmp %>% filter(season=="winter"),aes(y=NLL,colour=method,x=rep(names(winter),each=2))) + geom_point() + scale_color_manual(values=c("#009ADA", "#C11432"))

#geom_line(aes(x=method,y=NLL,group=c(cond_pollutant,season))
```
And for summer.

```{r}
ggplot(tmp %>% filter(season=="summer"),aes(y=NLL,colour=method,x=rep(names(winter),each=2))) + geom_point() + scale_color_manual(values=c("#009ADA", "#C11432"))

```

Compare AIC winter.

```{r}
ggplot(tmp %>% filter(season=="winter"),aes(y=AIC,colour=method,x=rep(names(winter),each=2))) + geom_point() + scale_color_manual(values=c("#009ADA", "#C11432"))
```


And summer.
```{r}
ggplot(tmp %>% filter(season=="summer"),aes(y=AIC,colour=method,x=rep(names(winter),each=2))) + geom_point() + scale_color_manual(values=c("#009ADA", "#C11432"))
```

Compare BIC winter.

```{r}
ggplot(tmp %>% filter(season=="winter"),aes(y=BIC,colour=method,x=rep(names(winter),each=2))) + geom_point() + scale_color_manual(values=c("#009ADA", "#C11432"))
```
And summer.

```{r}
ggplot(tmp %>% filter(season=="summer"),aes(y=BIC,colour=method,x=rep(names(winter),each=2))) + geom_point() + scale_color_manual(values=c("#009ADA", "#C11432"))
```
Maybe truncate?

## Print $PM_{10}>v$ fit for comparison.

```{r}
vcw1
summary(vcw1)
summary(vcw2)
```
And for summer.

```{r}
vcs1
summary(vcs1)
summary(vcs2)
```



