---
title: "Simulation examples"
output: html_document
date: "2025-03-13"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(GGally)
#library(MASS) # use dplyr::select to avoid function conflict
library(pracma) # incomplete gamma function, will be used later in cond.ext.model for modelling the residual margins
library(xtable)
library(LaplacesDemon)

source("../sample_distribution_helpers.R")
source("../likelihood_helpers.R")
source("../cond_model_helpers.R")

# set theme defaults to be black rectangle
theme_set(theme_bw())
theme_replace(
  panel.spacing = unit(2, "lines"),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  strip.background = element_blank(),
  panel.border = element_rect(colour = "black", fill = NA) )
```

## Two linked clusters

```{r}
set.seed(11)
N <- 5000
v <- 0.9
a <- 1/4
sims <- generate_Y(N=N) %>% link_log(dep=a) %>%
  link_norm(dep=3/4) %>% relocate(Y3,.before= Y1) %>% link_norm(dep=1/4) %>% link_log(dep=a) %>%
  apply(c(2),FUN=frechet_laplace_pit1) %>% as.data.frame() %>% relocate(Y3,.after = Y2)

# could save a large sample
# save(sims,file="../data_processed/simulation_examples.RData")
load("../data_processed/simulation_examples.RData")
```

Plot pairwise when $Y_1>v$.

```{r}
v <- 0.99
simsv <- sims %>% filter(Y1>quantile(Y1,v))
ggpairs(simsv)
```


# Fit a model to this example

Inference steps (Laplace margins):

1.  Estimate $\alpha$ and $\beta$.
2.  Calculate observed residuals.
3a.  Fit residual margins.
3b.  Fit a vine copula.

Simulation steps:

1.  Simulate $Y_1$ from a standard exponential plus the simulation threshold on a Laplace scale.
2.  Simulate residuals from a vine copula.
3.  Transform residuals from uniform to AGG margins.
4.  Calculate $\boldsymbol{Y}_{-1}$.
5.  Transform from Laplace to the original margins.

```{r}
j <- 1
# 1. estimate a and b
pe <- par_est(df=sims,v=v,given=j,margin = "Normal", method = "sequential2")
# 2. calculate observed residuals
obs_res <- observed_residuals(df = sims,given = j,v = v, a=pe$a,b=pe$b)
# 3a. estimate parameters for AGG residual margins
res_margin <- res_margin_par_est(obs_res=obs_res,method = "AGG")
# 3b. fit a vine copula
fit_res <- rvinecopulib::vinecop( (obs_res %>% apply(c(2),FUN=row_number))/(nrow(sims)*(1-v)+1),selcrit = "mbic")  
```

Compare residual margin densities of observed and simulated values.

```{r}
# ptmp <- rbind(obs_res %>% mutate(residuals="observed"),Zsim %>% mutate(residuals="simulated"))
# ggpairs(ptmp,aes(col=residuals)) 
```

Print estimated and true point estimates.

```{r}
# simulate a large sample for the empirical estimates
set.seed(15)
N <- 100000 # set lower before knitting as there is no need to save this object as lower N still gives useful estimates
v <- 0.9
# sims1 <- data.frame(Y1=qfrechet(p=runif(N,min=v,max=1))) %>% link_log(dep=3/4) %>%
#  link_norm(dep=3/4) %>% relocate(Y3,.before= Y1) %>% link_norm(dep=1/4) %>% link_log(dep=3/4) %>%
#   apply(c(2),FUN=frechet_laplace_pit1) %>% as.data.frame() %>% relocate(Y3,.after = Y2)

# calculate empirical estimates
u <- c(0.99,0.9925,0.995,0.9975,0.999)

# one of the clusters u
pu <- psim <- psim1 <- psim2 <- pu1 <- pu2 <- pemp <- pemp1 <- c()
uq <- sapply(1:length(u),function(i){unif_laplace_pit(u[i]) })
for (i in 1:length(u)) {
 pu[i] <- nrow(sims1 %>% filter(if_all(.cols=everything(),~.x>uq[i])))/nrow(sims1)*(1-v)
 pemp[i] <- nrow(sims %>% filter(if_all(.cols=everything(),~.x>uq[i])))/nrow(sims)
 pu1[i] <- nrow(sims1 %>% filter(if_all(.cols=all_of(1:3),~.x>uq[i])))/nrow(sims1)*(1-v)
 Nsim <- 1000
 pemp1[i] <- nrow(sims %>% filter(if_all(.cols=all_of(1:3),~.x>uq[i])))/nrow(sims)
 pu2[i] <- nrow(sims1 %>% filter(if_all(.cols=all_of(4:5),~.x>uq[i])))/nrow(sims1)*(1-v)
 psim[i] <- nrow(Gen_Y1 %>% filter(if_all(.cols=everything(),~.x>uq[i])))*(1-u[i])/Nsim
 psim1[i] <- nrow(Gen_Y1 %>% filter(if_all(.cols=all_of(1:3),~.x>uq[i])))*(1-u[i])/Nsim
}
```

Calculate bias(times 1000 for easier reading).

```{r}
(psim-pu)*1000
```

Convert into a table of results.

```{r}
tmp_results <- data.frame("Emp_P"=pemp*1000,"True_P" = pu*1000, "Est_P" = psim*1000, "Bias_P" = (psim-pu)*1000,"Emp_P1"=pemp1*1000,"True_P1" = pu1*1000, "Est_P1" = psim1*1000, "Bias_P1" = (psim1-pu1)*1000)
xtable(tmp_results,digits=4)
#,"True_P2" = pu2*1000, "Est_P2" = psim2*1000, "Bias_P2" = (psim2-pu2)*1000
```


Is this an improvement over other methods?

## Fit alternative methods for modelling the dependence of the residuals

We start with empirical residuals.

```{r, warning=FALSE}
set.seed(11)
Nsim <- 1000

# calculate the probabilities
psimemp <- psimemp1 <-   c()
uq <- sapply(1:length(u),function(i){unif_laplace_pit(u[i]) })
for (i in 1:length(u)) {
  Ysim <- simulate_cond_model(v_sim=u[i],param_estimates=pe,Nsim=Nsim,res_dist = "empirical")
 psimemp[i] <- nrow(Ysim %>% filter(if_all(.cols=everything(),~.x>uq[i])))*(1-u[i])/Nsim
 psimemp1[i] <- nrow(Ysim %>% filter(if_all(.cols=all_of(1:3),~.x>uq[i])))*(1-u[i])/Nsim

}

# could save large dataframes
# sims data, sims1 large data for true values
# Gen_Y1 model simulation, Ysim model simulation empirical residuals
# save(sims,sims1,file="../data_processed/simulation_examples.RData")
```


Table with other estimates.

```{r}
tmp_results <- data.frame("Emp_P"=pemp*1000,"True_P" = pu*1000, "Est_P" = psim*1000, "Bias_P" = (psim-pu)*1000,"Est_P_empres" = psimemp*1000, "Bias_P_empres" = (psimemp-pu)*1000)
xtable(tmp_results,digits=6)

# make 2 tables
tmp_results1 <- data.frame("Emp_P1"=pemp1*1000,"True_P1" = pu1*1000, "Est_P1" = psim1*1000, "Bias_P1" = (psim1-pu1)*1000,"Est_P1_empres" = psimemp1*1000, "Bias_P1_empres" = (psimemp1-pu1)*1000)
xtable(tmp_results1,digits=6)
#,"True_P2" = pu2*1000, "Est_P2" = psim2*1000, "Bias_P2" = (psim2-pu2)*1000

```
Compare the bias for different thresholds.

```{r}
Nrep <- 100
x <- sapply(1:Nrep,FUN=vine_bias,simplify=FALSE)
```

Plot estimates.

```{r}
tmp <- bind_rows(x, .id = "iteration") %>% mutate("threshold"=factor(rep(u,Nrep)))
tmpa <- tmp %>% dplyr::select(all_of(c(1,2,4,6,8))) %>% pivot_longer(cols=c(2,3,4)) %>% mutate(bias=value-pu) %>% mutate(name=factor(name))
p <- ggplot(tmpa) + geom_boxplot(aes(y=bias,fill=name)) + facet_wrap(facets = "threshold",nrow=1) +ylim(c(-0.0005,0.0005)) + scale_fill_discrete(name="Method",labels=c("empirical","CM vine copula","CM emp. residuals")) +   theme(axis.title.x=element_blank(),
    axis.text.x=element_blank(),
    axis.ticks.x=element_blank())
ggsave(p,filename="../../Documents/bias_compare100.png",width=10,height=3)

# repeat for cluster 1
tmp1 <- tmp %>% dplyr::select(all_of(c(1,3,5,7,8))) %>% pivot_longer(cols=c(2,3,4)) %>% mutate(bias=value-pu) %>% mutate(name=factor(name))
p <- ggplot(tmp1) + geom_boxplot(aes(y=bias,fill=name)) + facet_wrap(facets = "threshold",nrow=1) +ylim(c(-0.0005,0.0005)) + scale_fill_discrete(name="Method",labels=c("empirical","CM vine copula","CM emp. residuals")) +   theme(axis.title.x=element_blank(),
    axis.text.x=element_blank(),
    axis.ticks.x=element_blank())
ggsave(p,filename="../../Documents/bias1_compare100.png",width=10,height=3)

```

Also look at CI of the spread.

```{r}
tapply(tmpa$bias, tmpa$name, summary)
```

