---
title: "Simulation examples"
output: html_document
date: "2025-03-13"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(GGally)
#library(MASS) # use dplyr::select to avoid function conflict
library(pracma) # incomplete gamma function, will be used later in cond.ext.model for modelling the residual margins
library(texmex) # air pollution data
library(xtable)
source("../sample_distribution_helpers.R")
source("../likelihood_helpers.R")
source("../cond_model_helpers.R")

# set theme defaults to be black rectangle
theme_set(theme_bw())
theme_replace(
  panel.spacing = unit(2, "lines"),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  strip.background = element_blank(),
  panel.border = element_rect(colour = "black", fill = NA) )

```

## Two linked clusters

```{r}
set.seed(11)
N <- 500000
v <- 0.99
# sims <- generate_Y(N=N) %>% link_log(dep=3/4) %>%
#   link_norm(dep=3/4) %>% relocate(Y3,.before= Y1) %>% link_norm(dep=1/4) %>% link_log(dep=3/4) %>%
#   apply(c(1,2),FUN=frechet_laplace_pit) %>% as.data.frame() %>% relocate(Y3,.after = Y2)

# could save a large sample
# save(sims,file="../data_processed/simulation_examples.RData")
load("../data_processed/simulation_examples.RData")
```

Plot pairwise.

```{r}
#ggpairs(sims)
```

When $Y_1>v$.

```{r}
v <- 0.99
simsv <- sims %>% filter(Y1>quantile(Y1,v))
ggpairs(simsv)
```


# Fit a model to this example

Inference steps (Laplace margins):

1.  Estimate $\alpha$ and $\beta$.
2.  Calculate observed residuals.
3a.  Fit residual margins.
3b.  Fit a vine copula.

Simulation steps:

1.  Simulate $Y_1$ from a standard exponential plus the simulation threshold on a Laplace scale.
2.  Simulate residuals from a vine copula.
3.  Transform residuals from uniform to AGG margins.
4.  Calculate $\boldsymbol{Y}_{-1}$.
5.  Transform from Laplace to the original margins.

```{r}
j <- 1
# 1. estimate a and b
pe <- par_est(df=sims,v=v,given=j,margin = "AGG", method = "sequential2")
# 2. calculate observed residuals
obs_res <- observed_residuals(df = sims,given = j,v = v, a=pe$a,b=pe$b)
# 3a. get parameters for AGG residual margins
res_margin <- pe %>% dplyr::select(mu_agg,sigl,sigu,deltal,deltau)
# 3b. fit a vine copula
fit_res <- rvinecopulib::vinecop( (obs_res %>% apply(c(2),FUN=row_number))/(nrow(sims)*(1-v)+1))  
```

Simulation steps.

```{r, warning=FALSE}
# set.seed(11)
v_sim <- v
Nsim <- 100000
# # 1. simulate Y1
# U <- runif(Nsim)
# Y1_gen <- -log(2*(1-v_sim)) + rexp(Nsim)
# Gen_Y1 <- data.frame(Y1=Y1_gen)
# 
# # 2. simulate residuals from a vine copula
# Zsim_unif <- rvinecopulib::rvinecop(n=Nsim,vine=fit_res) %>% as.data.frame()
# 
# # 3. convert to original margins
# to_opt <- function(x) {
#   return( (F_AGG(x,theta=c(mu,sigl,sigu,deltal,deltau))-Zsim_unif[i,k])^2  )  
# }
# Zsim <- Zsim_unif
# for (k in 1:ncol(Zsim)) {
#   mu <- pe$mu_agg[k]
#   sigl <- pe$sigl[k]
#   sigu <- pe$sigu[k]
#   deltal <- pe$deltal[k]
#   deltau <- pe$deltau[k]
#   for (i in 1:nrow(Zsim)) {
#     Zsim[i,k] <- optim(fn=to_opt,par=1,method="Brent",lower=-10,upper=10)$par
#   }
# }
# 
# 
# # 4. calculate each Yj
# Y1 <- Gen_Y1$Y1
# Y2 <- pe$a[1]*Y1 + Y1^pe$b[1] *Zsim[,1]
# Y3 <-  pe$a[2]*Y1 + Y1^pe$b[2] *Zsim[,2]
# Y4 <- pe$a[3]*Y1 + Y1^pe$b[3] *Zsim[,3]
# Y5 <-  pe$a[4]*Y1 + Y1^pe$b[4] *Zsim[,4]
# res <- c(1:ncol(sims))[-j]
# Gen_Y1 <- Gen_Y1 %>% mutate(Y2=Y2,Y3=Y3,Y4=Y4,Y5=Y5) %>% mutate(sim="model")
# names(Gen_Y1) <- c(paste0("Y",j),paste0("Y",res[1]),paste0("Y",res[2]),paste0("Y",res[3]),paste0("Y",res[4]),"sim")
```

Compare residual margin densities of observed and simulated values.

```{r}
# ptmp <- rbind(obs_res %>% mutate(residuals="observed"),Zsim %>% mutate(residuals="simulated"))
# ggpairs(ptmp,aes(col=residuals)) 
```


Print estimated and true point estimates.

```{r}
# simulate a large sample for the empirical estimates
set.seed(15)
N <- 1000000 # set lower before knitting as there is no need to save this object as lower N still gives useful estimates
v <- 0.99
# sims1 <- data.frame(Y1=qfrechet(p=runif(N,min=v,max=1))) %>% link_log(dep=3/4) %>%
#   link_norm(dep=3/4) %>% relocate(Y3,.before= Y1) %>% link_norm(dep=1/4) %>% link_log(dep=3/4) %>%
#   apply(c(1,2),FUN=frechet_laplace_pit) %>% as.data.frame() %>% relocate(Y3,.after = Y2)

# calculate empirical estimates
psim <- pu <-  c()
u <- c(0.99,0.9925,0.995,0.9975,0.999)

# one of the clusters u
psim1 <- psim2 <- pu1 <- pu2 <- pemp <- pemp1 <-   c()
uq <- sapply(1:length(u),function(i){unif_laplace_pit(u[i]) })
for (i in 1:length(u)) {
 pu[i] <- nrow(sims1 %>% filter(if_all(.cols=everything(),~.x>uq[i])))/nrow(sims1)*(1-v)
 pemp[i] <- nrow(sims %>% filter(if_all(.cols=everything(),~.x>uq[i])))/nrow(sims)
 
 pu1[i] <- nrow(sims1 %>% filter(if_all(.cols=all_of(1:3),~.x>uq[i])))/nrow(sims1)*(1-v)

  pemp1[i] <- nrow(sims %>% filter(if_all(.cols=all_of(1:3),~.x>uq[i])))/nrow(sims)
 pu2[i] <- nrow(sims1 %>% filter(if_all(.cols=all_of(4:5),~.x>uq[i])))/nrow(sims1)*(1-v)
 psim[i] <- nrow(Gen_Y1 %>% filter(if_all(.cols=everything(),~.x>uq[i])))*(1-u[i])/Nsim
 psim1[i] <- nrow(Gen_Y1 %>% filter(if_all(.cols=all_of(1:3),~.x>uq[i])))*(1-u[i])/Nsim
 psim2[i] <- nrow(Gen_Y1 %>% filter(if_all(.cols=all_of(4:5),~.x>uq[i])))*(1-u[i])/Nsim 
}
pu*1000
psim*1000
```

Calculate bias(times 1000 for easier reading).

```{r}
(psim-pu)*1000
```

Convert into a table of results.

```{r}
tmp_results <- data.frame("Emp_P"=pemp*1000,"True_P" = pu*1000, "Est_P" = psim*1000, "Bias_P" = (psim-pu)*1000,"Emp_P1"=pemp1*1000,"True_P1" = pu1*1000, "Est_P1" = psim1*1000, "Bias_P1" = (psim1-pu1)*1000)
xtable(tmp_results,digits=4)
#,"True_P2" = pu2*1000, "Est_P2" = psim2*1000, "Bias_P2" = (psim2-pu2)*1000
```


Is this an improvement over other methods?

## Fit alternative methods for modelling the dependence of the residuals

We start with empirical residuals.


```{r, warning=FALSE}
set.seed(11)
v_sim <- v
Nsim <- 100000
# 1. simulate Y1
U <- runif(Nsim)
Y1sim <- -log(2*(1-v_sim)) + rexp(Nsim)
Ysim <- data.frame(Y1=Y1sim)

# 2. simulate residuals from a vector of observed residuals
Zsim <- obs_res[sample(x=1:nrow(obs_res),size=Nsim,replace=TRUE),]


# 4. calculate each Yj
Y1 <- Ysim$Y1
Y2 <- pe$a[1]*Y1 + Y1^pe$b[1] *Zsim[,1]
Y3 <-  pe$a[2]*Y1 + Y1^pe$b[2] *Zsim[,2]
Y4 <- pe$a[3]*Y1 + Y1^pe$b[3] *Zsim[,3]
Y5 <-  pe$a[4]*Y1 + Y1^pe$b[4] *Zsim[,4]
res <- c(1:ncol(sims))[-j]
Ysim <- Ysim %>% mutate(Y2=Y2,Y3=Y3,Y4=Y4,Y5=Y5) %>% mutate(sim="model")
names(Ysim) <- c(paste0("Y",j),paste0("Y",res[1]),paste0("Y",res[2]),paste0("Y",res[3]),paste0("Y",res[4]),"sim")

# calculate the probabilities
psimemp <- psimemp1 <-   c()
uq <- sapply(1:length(u),function(i){unif_laplace_pit(u[i]) })
for (i in 1:length(u)) {
 psimemp[i] <- nrow(Ysim %>% filter(if_all(.cols=everything(),~.x>uq[i])))*(1-u[i])/Nsim
 psimemp1[i] <- nrow(Ysim %>% filter(if_all(.cols=all_of(1:3),~.x>uq[i])))*(1-u[i])/Nsim

}

# could save large dataframes
# sims data, sims1 large data for true values
# Gen_Y1 model simulation, Ysim model simulation empirical residuals
# save(sims,sims1,Gen_Y1,Ysim,file="../data_processed/simulation_examples.RData")
```


Table with other estimates.

```{r}
tmp_results <- data.frame("Emp_P"=pemp*1000,"True_P" = pu*1000, "Est_P" = psim*1000, "Bias_P" = (psim-pu)*1000,"Est_P_empres" = psimemp*1000, "Bias_P_empres" = (psimemp-pu)*1000)
xtable(tmp_results,digits=6)

# make 2 tables
tmp_results1 <- data.frame("Emp_P1"=pemp1*1000,"True_P1" = pu1*1000, "Est_P1" = psim1*1000, "Bias_P1" = (psim1-pu1)*1000,"Est_P1_empres" = psimemp1*1000, "Bias_P1_empres" = (psimemp1-pu1)*1000)
xtable(tmp_results1,digits=6)
#,"True_P2" = pu2*1000, "Est_P2" = psim2*1000, "Bias_P2" = (psim2-pu2)*1000

```

