---
title: "Vine structure comparison"
output: html_document
date: "2025-03-18"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(VineCopula)
library(rvinecopulib)
library(tidyverse)
library(texmex)
```

```{r}
v <- 0.7
# transform to Laplace margins
winter_lap <- as.data.frame((winter %>% apply(c(2),FUN=row_number))/(nrow(winter)+1)) %>% apply(c(1,2),FUN=unif_laplace_pit) %>% as.data.frame()
summer_lap <- as.data.frame((summer %>% apply(c(2),FUN=row_number))/(nrow(summer)+1)) %>% apply(c(1,2),FUN=unif_laplace_pit) %>% as.data.frame()
pew <-  par_est(df=winter_lap,v=v,given=j,margin = "AGG", method = "sequential2")
obsresw <- (observed_residuals(df = winter_lap,given = 1,v = v,a = pew$a,b=pew$b) %>% apply(c(2),FUN=row_number))/(nrow(winter_lap)*(1-v)+1)

pes <-  par_est(df=summer_lap,v=v,given=j,margin = "AGG", method = "sequential2")
obsress <- (observed_residuals(df = summer_lap,given = 1,v = v,a = pes$a,b=pes$b) %>% apply(c(2),FUN=row_number))/(nrow(summer_lap)*(1-v)+1)

vcw1 <- rvinecopulib::vinecop(data=obsresw)
vcw1
summary(vc1)
# compare with VineCopula fit
vcw2 <- VineCopula::RVineStructureSelect(data=obsresw)
summary(vcw2)

vcs1 <- rvinecopulib::vinecop(data=obsress)
vcs1
summary(vcs1)
# compare with VineCopula fit
vcs2 <- VineCopula::RVineStructureSelect(data=obsress)
summary(vcs2)

```
```

## Impose tree structure and copula families


## Compare imposing one link at a time
```{r}
summary(vc2)

bivc <- VineCopula::RVineStructureSelect(data=obsresw[,1:2])
RVineSeqEst(data=obsress[,1:2],RVM=bivc)

# look at other way around
bivc <- VineCopula::RVineStructureSelect(data=obsress[,1:2])
RVineSeqEst(data=obsresw[,1:2],RVM=bivc)


```

Try another pair

```{r}
summary(vc2)

bivc <- VineCopula::RVineStructureSelect(data=obsresw[,3:4])
RVineSeqEst(data=obsress[,3:4],RVM=bivc)

# look at other way around
bivc <- VineCopula::RVineStructureSelect(data=obsress[,3:4])
RVineSeqEst(data=obsresw[,3:4],RVM=bivc)


```

