---
title: "Vine structure comparison"
output: html_document
date: "2025-03-18"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(VineCopula)
library(rvinecopulib)
library(tidyverse)
library(texmex)
#library(ggraph) # for plotting vines
source("../sample_distribution_helpers.R")
source("../likelihood_helpers.R")
source("../cond_model_helpers.R")

# set theme defaults to be black rectangle
theme_set(theme_bw())
theme_replace(
  panel.spacing = unit(2, "lines"),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  strip.background = element_blank(),
  panel.border = element_rect(colour = "black", fill = NA) )

```

```{r}
v <- 0.7
# transform to Laplace margins
winter_lap <- as.data.frame((winter %>% apply(c(2),FUN=row_number))/(nrow(winter)+1)) %>% apply(c(1,2),FUN=unif_laplace_pit) %>% as.data.frame()
summer_lap <- as.data.frame((summer %>% apply(c(2),FUN=row_number))/(nrow(summer)+1)) %>% apply(c(1,2),FUN=unif_laplace_pit) %>% as.data.frame()

# calculate AIC of the first tree
T1_AIC <- function(x) {
  y <- summary(x)
-2*sum(y$loglik[1:3])+2*length(as.numeric(unlist(y$parameters[1:3])))
}

# calculate log-likelihood of the first tree
T1l<- function(x) {
  y <- summary(x)
-2*sum(y$loglik[1:3])
}

# return a row of 2nll and AIC of the first tree of winter and summer fits
nllAICsw <- function(vcw,vcs) {
  x <- c(T1_AIC(vcw),T1l(vcw),T1_AIC(vcs),T1l(vcs),T1_AIC(vcw)+T1_AIC(vcs),T1l(vcw)+T1l(vcs))
  return(round(x,2))
}
```

Compare the 5 methods conditioning on each of the variables.

```{r}
pew <-  par_est(df=winter_lap,v=v,given=j,margin = "Normal", method = "sequential2")
obsresw <- (observed_residuals(df = winter_lap,given = j,v = v,a = pew$a,b=pew$b) %>% apply(c(2),FUN=row_number))/(nrow(winter_lap)*(1-v)+1)

pes <-  par_est(df=summer_lap,v=v,given=j,margin = "Normal", method = "sequential2")
obsress <- (observed_residuals(df = summer_lap,given = j,v = v,a = pes$a,b=pes$b) %>% apply(c(2),FUN=row_number))/(nrow(summer_lap)*(1-v)+1)

# 1. separate estimates for winter and summer residuals
vcw1 <- rvinecopulib::vinecop(data=obsresw,selcrit = "mbic")
vcs1 <- rvinecopulib::vinecop(data=obsress,selcrit = "aic")
vcw1
summary(vcw1)

m1 <- nllAICsw(vcw=vcw1,vcs=vcs1)

# 2. winter with imposed summer structure
vcw3 <- rvinecopulib::vinecop(data=obsresw,structure=get_structure(vcs1))
T1_AIC(vcw3)
vcs3 <- rvinecopulib::vinecop(data=obsress,structure=get_structure(vcs1))
T1_AIC(vcw3) + T1_AIC(vcs3)
T1l(vcw3) + T1l(vcs3)

# 4. joined data
rvine_join <- rvinecopulib::vinecop(data=rbind(obsresw,obsress)) 
T1_AIC(rvine_join)
T1l(rvine_join)

# 6. imposed tree structure and family from summer vinecop

```


## Impose tree structure and copula families


## Compare imposing one link at a time

Try the residuals of $NO_2$ and $NO$.

```{r}

bivc <- VineCopula::RVineStructureSelect(data=obsresw[,1:2])
RVineSeqEst(data=obsress[,1:2],RVM=bivc)

# look at other way around
bivc <- VineCopula::RVineStructureSelect(data=obsress[,1:2])
RVineSeqEst(data=obsresw[,1:2],RVM=bivc)


```

Try another pair of $SO_2$ and $PM_{10}$ residuals.

```{r}
bivc <- VineCopula::RVineStructureSelect(data=obsresw[,3:4])
RVineSeqEst(data=obsress[,3:4],RVM=bivc)

# look at other way around
bivc <- VineCopula::RVineStructureSelect(data=obsress[,3:4])
RVineSeqEst(data=obsresw[,3:4],RVM=bivc)

```

## Try all possible bivariate copulas for a given link

```{r}
index_cop <- c(0:10,13,16,18,20,23,24,26:30,33,34,36:40,104,114,124,134,204,214,224)
x <- list()
tmp <- obsresw[,1:2]
tmp1 <- obsress[,1:2]

for (i in 1:length(index_cop)) {
  tryCatch({
x[[i]] <-  BiCopEst(u1=as.numeric(tmp[,1]),u2=as.numeric(tmp[,2]),family=i)$AIC
  },error=function(e){print(NA)})
}
```

Try the Shiny app for comparing copula fits.

```{r}
# BiCopCompare(u1=obsresw[,1],u2=obsresw[,2])
# 
# # Try another pair of SO2 and PM10
# BiCopCompare(u1=obsresw[,3],u2=obsresw[,4])
# 
# 
# BiCopCompare(u1=obsresw[,2],u2=obsresw[,3])
```

## Compare all possible copulas for a given link

```{r, echo=FALSE}
d <- 4
tmpdf <- tmpdfs <-  data.frame(pair_index=as.numeric(),u1=as.numeric(),u2=as.numeric(),AIC=as.numeric(),familyname= as.character(),family=as.numeric(),par=as.numeric(),par2=as.numeric())
# write a loop for all pairs in the first tree
for (m in 1:(d-1)) {
  index_var1 <- vcw2$Matrix[m,m]
  index_var2 <- vcw2$Matrix[4,m]
  index_var1s <- vcs2$Matrix[m,m]
  index_var2s <- vcs2$Matrix[4,m]
  # calculate all possible copulas
  for (i in 1:length(index_cop)) {
    x <- BiCopEst(u1=as.numeric(obsresw[,index_var1]),u2=as.numeric(obsresw[,index_var2]),family=index_cop[i])
    tmpdf <- rbind(tmpdf,data.frame(pair_index=paste0(names(winter)[index_var1+1],"_",names(winter)[index_var2+1]),u1=index_var1,u2=index_var2,AIC=x$AIC,familyname=x$familyname,family=x$family,par=x$par,par2=x$par2))
   y <- BiCopEst(u1=as.numeric(obsress[,index_var1s]),u2=as.numeric(obsress[,index_var2s]),family=index_cop[i])
    tmpdfs <- rbind(tmpdfs,data.frame(pair_index=paste0(names(winter)[index_var1s+1],"_",names(winter)[index_var2s+1]),u1=index_var1s,u2=index_var2s,AIC=y$AIC,familyname=y$familyname,family=y$family,par=y$par,par2=y$par2))

  }
}
```

```{r}
# plot for each pair
ggplot(tmpdf) + geom_bar(aes(x=familyname,y=sort(AIC)),stat="identity") + facet_wrap(~pair_index)
```
What are the top 5?

```{r}
tmpdftop5 <- tmpdf %>% group_by(pair_index) %>% arrange(-desc(AIC)) %>% slice(1:5)
tmpdftop5 %>% ggplot() + geom_bar(aes(x=factor(family),y=sort(AIC)),stat="identity") + facet_wrap(~pair_index,scales="free_x")
tmpdftop5s <- tmpdfs %>% group_by(pair_index) %>% arrange(-desc(AIC)) %>% slice(1:5)
tmpdftop5s %>% ggplot() + geom_bar(aes(x=factor(family),y=sort(AIC)),stat="identity",col="#C11432") + facet_wrap(~pair_index,scales="free_x")
```

```{r}
tmpdftop5
```

```{r}
tmpdftop5s
```


Which one gives the lowest sum (winter + summer) of the AIC for $NO_2$ and $NO$ pair of residuals?

```{r}
sAIC <- tmpdfs %>% filter(pair_index=="SO2_NO") %>% pull(AIC)
tmpdf1 <- tmpdf %>% filter(pair_index=="SO2_NO") %>% mutate(AICsum=AIC+sAIC)
tmpdf1 <- tmpdf1 %>% arrange(-desc(AICsum)) %>% relocate(AICsum,.after=AIC)
tmpdf1
```
Compare the top 5 rows for winter and summer.

```{r}
tmpdf1[1:5,]
tmpdfs %>% filter(pair_index=="SO2_NO", family %in% tmpdf1[1:5,7])
```

To generalise this, we could look at all the pair that are in either winter or summer tree 1 structure.

Could we maximise sum of $|\tau_{winter} + \tau_{summer}|$ to determine the pairs?

```{r,echo=FALSE}
pairs <- expand.grid(1:4,1:4) %>% filter(Var1>Var2)
tmpdf <- tmpdfs <-  data.frame(pair_index=as.numeric(),u1=as.numeric(),u2=as.numeric(),tau=as.numeric(),AIC=as.numeric(),familyname= as.character(),family=as.numeric(),par=as.numeric(),par2=as.numeric())
# write a loop for all pairs in the first tree
for (m in 1:nrow(pairs)) {
  index_var1 <- pairs[m,1]
  index_var2 <- pairs[m,2]
  # calculate all possible copulas
  for (i in 1:length(index_cop)) {
    x <- BiCopEst(u1=as.numeric(obsresw[,index_var1]),u2=as.numeric(obsresw[,index_var2]),family=index_cop[i])
    tmpdf <- rbind(tmpdf,data.frame(pair_index=paste0(names(winter)[index_var1+1],"_",names(winter)[index_var2+1]),u1=index_var1,u2=index_var2,tau=x$tau,AIC=x$AIC,familyname=x$familyname,family=x$family,par=x$par,par2=x$par2))
   y <- BiCopEst(u1=as.numeric(obsress[,index_var1]),u2=as.numeric(obsress[,index_var2]),family=index_cop[i])
    tmpdfs <- rbind(tmpdfs,data.frame(pair_index=paste0(names(winter)[index_var1+1],"_",names(winter)[index_var2+1]),u1=index_var1,u2=index_var2,tau=y$tau,AIC=y$AIC,familyname=y$familyname,family=y$family,par=y$par,par2=y$par2))

  }
}

```
Look at top 5 for absolute tau.

```{r}
tmpdftop5 <- tmpdf %>% group_by(pair_index) %>% arrange(desc(abs(tau))) %>% slice(1:5)
tmpdftop5 %>% ggplot() + geom_bar(aes(x=factor(family),y=sort(abs(tau))),stat="identity") + facet_wrap(~pair_index,scales="free_x")
tmpdftop5s <- tmpdfs %>% group_by(pair_index) %>% arrange(desc(abs(tau))) %>% slice(1:5)
tmpdftop5s %>% ggplot() + geom_bar(aes(x=factor(family),y=sort(abs(tau))),stat="identity",col="#C11432") + facet_wrap(~pair_index,scales="free_x")
```

Why does tau depend on the family?


```{r}
tmpdf <- tmpdf %>% mutate(taus=tmpdfs$tau) %>% rowid_to_column()
tmpdf1 <- tmpdf %>% mutate(tausum=abs(taus+tau))
tmpdf1 <- tmpdf1 %>% group_by(pair_index) %>%  arrange(desc(tausum)) %>% relocate(tausum,taus,.after=tau) %>% slice(1)
tmpdf1
tmpdfs <- tmpdfs %>% rowid_to_column()
t <- tmpdf1$rowid
tmpdfs[t,]
```
Compare the top 5 rows for winter and summer.

```{r}
tmpdf5 <- tmpdf %>% mutate(tausum=abs(taus+tau))
tmpdf5 <- tmpdf5 %>% group_by(pair_index) %>%  arrange(desc(tausum)) %>% relocate(tausum,taus,.after=tau) %>% slice(1:5)
tmpdf5
t <- tmpdf5$rowid
tmpdfs[t,]
```


## Moving into conditional trees

Try and compare for a given conditional pair. These are not common between summer and winter, try just 41|2, that is $PM_{10}NO_2|NO$ residuals.

```{r}
# 4 given 2
u1cond <- BiCopHfunc(u1=obsresw[,4],u2=obsresw[,2],family = 104,par=1.74,par2=0.47)$hfunc2
# 1 given 2
u2cond <- BiCopHfunc(u1=obsresw[,1],u2=obsresw[,2],family = 20,par=6,par2=0.8)$hfunc2
  
BiCopSelect(u1=u1cond,u2=u2cond)  
# compare
vcw2

BiCopHfunc(u1=u1cond,u2=u2cond,)
```
## Automate into a loop

The idea is to repeat the algorithm but each step picks the best structure overall for the two datasets.

```{r}
fit_copula_two_dataset <- function(data1=obsresw,data2=obsress) {
  d <- ncol(obsresw)
  # 1. fit a common tree structure
  rvine_winter <- rvinecopulib::vinecop(data=obsresw, selcrit = "mbicv") 
  rvine_summer <- rvinecopulib::vinecop(data=obsress, selcrit = "mbicv") 
  rvine_join <- rvinecopulib::vinecop(data=rbind(obsresw,obsress), selcrit = "mbicv") 
  get_matrix(rvine_join)
  if (identical(get_matrix(rvine_winter),get_matrix(rvine_join)) | identical(get_matrix(rvine_summer),get_matrix(rvine_join)) ) {
    # impose common structure
    tree_struc <- get_matrix(rvine_join)
  # calculate best fitting overall bivariate copula for each pair in T1
<<<<<<< HEAD
tmpdfw <- tmpdfs <-  data.frame(pair_index=numeric(),u1=numeric(),u2=numeric(),tau=numeric(),AIC=numeric(),ll=numeric(),familyname= character(),family=numeric(),par=numeric(),par2=numeric())
=======
tmpdfw <- tmpdfs <-  data.frame(tree=numeric(),edge=numeric(),pair_index=numeric(),u1=numeric(),u2=as.numeric(),tau=as.numeric(),AIC=as.numeric(),familyname= as.character(),family=as.numeric(),par=as.numeric(),par2=as.numeric())
>>>>>>> febe180f0778add8cc2d24d356b3a95de2ba6825
# set up dataframe to store family
familydf <- summary(rvine_join)[,c(1:4,6:7)] %>% mutate(joint_family=NA,joint_family_index=NA)
# write a loop for all pairs in the first tree
for (m in 1:(d-1)) {
  index_var1 <- familydf[m,3][[1]][1]
  index_var2 <- familydf[m,3][[1]][2]
  # calculate all possible copulas
  for (i in 1:length(index_cop)) {
    x <- BiCopEst(u1=as.numeric(obsresw[,index_var1]),u2=as.numeric(obsresw[,index_var2]),family=index_cop[i])
<<<<<<< HEAD
    tmpdfw <- rbind(tmpdfw,data.frame(pair_index=paste0(names(winter)[index_var1+1],"_",names(winter)[index_var2+1]),u1=index_var1,u2=index_var2,tau=x$tau,AIC=x$AIC,ll=x$logLik,familyname=x$familyname,family=x$family,par=x$par,par2=x$par2))
   y <- BiCopEst(u1=as.numeric(obsress[,index_var1]),u2=as.numeric(obsress[,index_var2]),family=index_cop[i])
    tmpdfs <- rbind(tmpdfs,data.frame(pair_index=paste0(names(winter)[index_var1+1],"_",names(winter)[index_var2+1]),u1=index_var1,u2=index_var2,tau=y$tau,AIC=y$AIC,ll=y$logLik,familyname=y$familyname,family=y$family,par=y$par,par2=y$par2))
  }
  
  }
tmpdfw <- tmpdfw %>% mutate(AICs=tmpdfs$AIC,lls=tmpdfs$ll) %>% rowid_to_column()
tmpdf1 <- tmpdfw %>% mutate(AICsum=(AICs+AIC))
tmpdf1 <- tmpdf1 %>% group_by(pair_index) %>%  arrange(-desc(AICsum)) %>% relocate(AICsum,AICs,.after=AIC) %>% slice(1)

tmpdf5w <- tmpdfw %>% mutate(AICsum=(AICs+AIC))
tmpdf5w <- tmpdf5w %>% group_by(pair_index) %>%  arrange(-desc(AICsum)) %>% relocate(AICsum,AICs,.after=AIC) %>% slice(1:5)
t <- tmpdf5w$rowid
tmpdf5s <- tmpdfs[t,]
# save copula families in first tree in a matrix
familydf[1:(d-1),c(7:8)] <- tmpdf1[1:(d-1),c(10:11)]
=======
    tmpdfw <- rbind(tmpdfw,data.frame(tree=1,edge=m,pair_index=paste0(names(winter)[index_var1+1],"_",names(winter)[index_var2+1]),u1=index_var1,u2=index_var2,tau=x$tau,AIC=x$AIC,familyname=x$familyname,family=x$family,par=x$par,par2=x$par2))
   y <- BiCopEst(u1=as.numeric(obsress[,index_var1]),u2=as.numeric(obsress[,index_var2]),family=index_cop[i])
    tmpdfs <- rbind(tmpdfs,data.frame(tree=1,edge=m,pair_index=paste0(names(winter)[index_var1+1],"_",names(winter)[index_var2+1]),u1=index_var1,u2=index_var2,tau=y$tau,AIC=y$AIC,familyname=y$familyname,family=y$family,par=y$par,par2=y$par2))
  }

  }
tmpdfw <- tmpdfw %>% mutate(taus=tmpdfs$tau) %>% rowid_to_column()
tmpdf1 <- tmpdfw %>% mutate(tausum=abs(taus+tau))
tmpdf1 <- tmpdf1 %>% group_by(edge) %>%  arrange(desc(tausum)) %>% relocate(tausum,taus,.after=tau) %>% slice(1) %>% arrange(-desc(edge))

tmpdf5w <- tmpdfw %>% mutate(tausum=abs(taus+tau))
tmpdf5w <- tmpdf5w %>% group_by(edge) %>%  arrange(desc(tausum)) %>% relocate(tausum,taus,.after=tau) %>% slice(1:5)
t <- tmpdf5w$rowid
tmpdf5s <- tmpdfs[t,]
# save copula families in first tree in a matrix
familydf[1:(d-1),c(7:8)] <- tmpdf1[1:(d-1),c(11:12)]
>>>>>>> febe180f0778add8cc2d24d356b3a95de2ba6825

  # find best fitting overall cond.copula in T2,...
u1cond <- u2cond <- list()
tmpdfw <- tmpdfs <-  data.frame(tree=numeric(),edge=numeric(),pair_index=numeric(),u1=numeric(),u2=as.numeric(),tau=as.numeric(),AIC=as.numeric(),familyname= as.character(),family=as.numeric(),par=as.numeric(),par2=as.numeric())

#   for (t in 2:(d-2)) {
# for (j in 1:(d-2)) {
#   # select edge 1
#  ucond <-  familydf %>% filter(tree==t,edge==j) 
#  u11 <- obsresw[,(ucond %>% pull(conditioned))[[1]][1]]
# u1cond <- obsresw[,(ucond %>% pull(conditioning))[[1]][1]]
# u12 <- obsresw[,(ucond %>% pull(conditioned))[[1]][2]]
# 
# # find which family
# f1 <- f2 <- c()
# for (l in 1:(d-1)) {
#  f1[l] <-  identical(sort((familydf %>% filter(tree==1))$conditioned[[l]]),sort(c((ucond %>% pull(conditioned))[[1]][1],(ucond %>% pull(conditioning))[[1]][1])))
#  f2[l] <- identical(sort((familydf %>% filter(tree==1))$conditioned[[l]]), sort(c((ucond %>% pull(conditioned))[[1]][2],(ucond %>% pull(conditioning))[[1]][1])))
# }
# #family1 <- (familydf %>% filter(tree==1))[familydf$conditioned==sort(c(1,3)),]
# family1 <- (familydf %>% filter(tree==1))$joint_family_index[f1]
# 
# family2 <- (familydf %>% filter(tree==1))$joint_family_index[f2]
# parf1 <- 
# par2f1
# 
# u1cond[[j]] <- BiCopHfunc(u1=u11,u2=u1cond,family = family1,par=1.74,par2=0.47)$hfunc2
# # 1 given 2
# u2cond[[j]] <- BiCopHfunc(u1=u12,u2=u1cond,family = family2,par=6,par2=0.8)$hfunc2
# 
# #repeat for summer
# 
# 
#  for (i in 1:length(index_cop)) {
#     x <- BiCopEst(u1=as.numeric(u1cond[[j]]),u2=as.numeric(u2cond[[j]]),family=index_cop[i])
#     tmpdfw <- rbind(tmpdfw,data.frame(tree=t,edge=j,pair_index=paste0(names(winter)[index_var1+1],"_",names(winter)[index_var2+1]),u1=index_var1,u2=index_var2,tau=x$tau,AIC=x$AIC,familyname=x$familyname,family=x$family,par=x$par,par2=x$par2))
#    y <- BiCopEst(u1=as.numeric(obsress[,index_var1]),u2=as.numeric(obsress[,index_var2]),family=index_cop[i])
#     tmpdfs <- rbind(tmpdfs,data.frame(tree=t,edge=j,pair_index=paste0(names(winter)[index_var1+1],"_",names(winter)[index_var2+1]),u1=index_var1,u2=index_var2,tau=y$tau,AIC=y$AIC,familyname=y$familyname,family=y$family,par=y$par,par2=y$par2))
#   }
# 
# 
# }
#     tmpdfw <- tmpdfw %>% mutate(taus=tmpdfs$tau) %>% rowid_to_column()
# tmpdf2 <- tmpdfw %>% mutate(tausum=abs(taus+tau))
# tmpdf2 <- tmpdf2 %>% group_by(edge) %>%  arrange(desc(tausum)) %>% relocate(tausum,taus,.after=tau) %>% slice(1) %>% arrange(-desc(edge))
# 
# tmpdf5w2 <- tmpdfw %>% mutate(tausum=abs(taus+tau))
# tmpdf5w2 <- tmpdf5w2 %>% group_by(edge) %>%  arrange(desc(tausum)) %>% relocate(tausum,taus,.after=tau) %>% slice(1:5)
# t <- tmpdf5w2$rowid
# tmpdf5s2 <- tmpdfs[t,]
# # save copula families in first tree in a matrix
# familydf[4:5,c(7:8)] <- tmpdf2[1:(d-2),c(11:12)]
# 
# }

   # end for loop for conditional copula trees
return(list(familydf,tmpdf1,tmpdf5w,tmpdf5s))
  } # end if tree structure match
  else {
    # if Vuong test rejects, only copula pairs in common between two datasets can be compared in step 2 and 3
    print("No match between joined tree structure and summer/winter data.")
  }
  
}
```

Try.

```{r}
# summarise analysis to report AIC and NLL values
j <- 1
pew <-  par_est(df=winter_lap,v=v,given=j,margin = "Normal", method = "sequential2")
obsresw <- (observed_residuals(df = winter_lap,given = j,v = v,a = pew$a,b=pew$b) %>% apply(c(2),FUN=row_number))/(nrow(winter_lap)*(1-v)+1)

pes <-  par_est(df=summer_lap,v=v,given=j,margin = "Normal", method = "sequential2")
obsress <- (observed_residuals(df = summer_lap,given = j,v = v,a = pes$a,b=pes$b) %>% apply(c(2),FUN=row_number))/(nrow(summer_lap)*(1-v)+1)

# 1. separate fits to winter and summer data
vcw1 <- rvinecopulib::vinecop(data=obsresw)
summary(vcw1)
# compare with VineCopula fit
vcs1 <- rvinecopulib::vinecop(data=obsress)
summary(vcs1)
# sum
T1_AIC(vcs1) + T1_AIC(vcw1)
# calculate log-likelihood
T1l(vcw1) + T1l (vcs1)

# 2. imposed summer structure
# winter with imposed summer structure
vcw3 <- rvinecopulib::vinecop(data=obsresw,structure=get_structure(vcs1))
# test for summer structure also
vcs3 <- rvinecopulib::vinecop(data=obsress,structure=get_structure(vcs1))
T1_AIC(vcw3) + T1_AIC(vcs3)
T1l(vcw3) + T1l(vcs3)

fc <- fit_copula_two_dataset()
sum(fc[[2]]$AIC)+ sum(fc[[2]]$AICs)
fc

# 4. joined structure
 rvine_join <- rvinecopulib::vinecop(data=rbind(obsresw,obsress), selcrit = "mbicv") 
summary(rvine_join)
T1_AIC(rvine_join)
T1l(rvine_join)
 
-2*sum(fc[[2]]$ll)
# impose tree and family of joined fit
rvw5 <- rvinecopulib::vinecop(data=obsresw,structure = get_matrix(rvine_join),family_set = c("gaussian","clayton"),allow_rotations = FALSE)
rvs5 <- rvinecopulib::vinecop(data=obsress,structure = get_matrix(rvine_join),family_set = c("gaussian","clayton"),,allow_rotations = FALSE)
summary(rvw5)
summary(rvs5)
T1_AIC(rvw5) + T1_AIC(rvs5)
T1l(rvw5) + T1l(rvs5)

```


### Additional notes on fitting vine copulas

The default selection criterion is AIC, but possible alternatives are BIC, mBIC and mBICv. Furthermore, mBICv has a parameter $\psi_0$ for prior probability of  (lower $\psi_0$ tends to truncate earlier since non-independence copulas are more likely).

Question for Ingrid: One argument of fitting vine copula in `rvinecopulib` package is `presel`: whether the family set should be thinned out according to the symmetry characteristics of the data.
